### 2020-03-10 paper/main.tex
initial commit

\documentclass{turabian-thesis}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{pslatex}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{paralist}
\usepackage[utf8]{inputenc}
\usepackage{multicol} 
\usepackage[hidelinks]{hyperref}
\usepackage{nameref}

\DeclareUnicodeCharacter{2215}{\textminus}
\usepackage{algorithm}
\usepackage{algpseudocode}
\algdef{SE}[DOWHILE]{Do}{DoWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{claim}{Claim}
\newtheorem{definition}{Definition}
\newtheorem{problem}{Problem}

\newcounter{case}[section]
\newenvironment{case}[1][]{\refstepcounter{case}\par\medskip
   \textbf{Case~\thecase. #1} \rmfamily}{\medskip}

\newcounter{subcase}[case]
\newenvironment{subcase}[1][]{\refstepcounter{subcase}\par\medskip
   \textbf{Case~\thecase.\thesubcase. #1} \rmfamily}{\medskip}

\begin{document}

\frontmatter
\thispagestyle{empty}

\begin{center}

   \textbf{hybrid model}

   \vspace*{\baselineskip}

   A THESIS \\
   Presented to the Department of Computer Science and Computer Engineering \\
   California State University, Long Beach

   \vspace*{\baselineskip}

   In Partial Fulfillment \\
   of the Requirements for the Degree \\
   Master of Science in Computer Science \\
   Option in Computer Science

   \vspace*{\baselineskip}

   Committee Members: \\

   \vspace*{\baselineskip}

   By Ruben Rosales \\
   May 2020
\end{center}

\pagebreak

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\tableofcontents
\listofillustrations

\mainmatter

\chapter{Introduction}
In this thesis, we study interpretable machine learning as applied to complex-valued radio signals. Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification. These methods, however, fall short of allowing users to visualize patterns within their dataset and . To address this issue, we propose a hybrid interpretable model that can be extended to any time series dataset. Our model consists of two models. The first is an ensemble method of classifiers that functions as our black box method. The second is a white-box model that encodes time series as images for each signal and outputs all images that have been correctly classified.

We study the adaption of deep learning models to complex-valued radio signals. We compare the deep learning models such as CNN’s have become ever-increasingly powerful and widely used but their complex nature makes it difficult for users to understand what is going on so as black box models get more popular the need for an interpretable accomplice is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.

The growth of mobile devices and demand for wireless data has created a need for  High quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.  however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments. [7] [6] [9]
This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically

Building upon successful strategies from image and voice recognition domains in machine learning, we demonstrate an approach in the radio using Convolutional Neural Networks (CNNs) and Deep Neural Networks (DNNs) which offers flexibility to learn features across a wide range of tasks and demonstrates improved classification accuracy against current day approaches.  Existing methods, however, fall short of allowing users to visualize patterns within their dataset. To address this issue, we propose a hybrid interpretable model that can be extended to any time series dataset. Our model consists of two models. The first is an ensemble method of classifiers that functions as our black box method. The second is a white-box model that encodes time series as images for each signal and outputs all images that have been correctly classified.
\section{Contributions}

\section{Models}
Our black box method is composed of two types of neural networks, Convolutional Neural Networks and Long Short Term Memory networks.

CNN
Our CNN has an architecture consisting of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X). We found no improvements in performance when we attempted to make it deeper by adding additional convolutional layers as well as skip connections.

LSTM
We wanted to keep our LSTM network as small as possible, for X purposes, so we went with two LSTM layers of 128 hidden states and a dropout rate of 4/10  followed by a dense layer of 128 connections. We attempted different state sizes but found 128 to be the smallest number with the best consistent performance.

\section{black box}
Our model is constructed of four different architectures, 2 LSTM and 2 CNN’s, that are concatenated at the dense layer. The four features we use are amplitude and phase of each level of decomposition for HV and VV. We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other. Wavelet decomposition produces a complex value but that is unusable in a standard neural network due to loss of information unless we move to a complex LSTM but as stated in {} there isn’t a significant increase in accuracy.

\section{white box}
Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method. We then pass those images through a CNN and those that are correctly classified are shown to the user.

In order to classify those images we created a CNN for it specifically and avoided using our existing CNN in order to make it deeper and increase performance when images aren’t that classifiable. We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it. We were able to achieve 60% accuracy on our dataset.

\section{Notation}

\section{Layout}

\chapter{Problem Statement}

\chapter{Related Work}

\chapter{Markov Transition Fields}
% We propose a framework similar to [] for encoding dynamical transition statistics,
% but we continue their work by considering ith order Markov transition probabilities.
% Given a time series X, we decompose its magnitude axis into two separate
% properties by representing it as a polar coordinate, Xangular and Xradial. We then
% identify the Q quantile bins for both properties and the temporal values Xtime
% and assign each x ∈ Xangular, Xradial, Xtime to its corresponding bin qj (j in
% [1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
% Wradial, Wtime, by counting transitions among quantile bins in the manner of a
% ith order Markov chain along the time axis.

% W does not take into account the temporal axis so to prevent any information
% loss we construct a Markov Transition Field (MTF) for each W. The MTF
% denotes the probability of transitioning from qi to qj for each x ∈ X. This,
% in turn, allows us to consider the transition probability on the magnitude and
% temporal axis.

% As described in [] the MTF encodes the multi-span transition probabilities
% of the time series, but given that we have three different M’s we modify their
% approach and consider each M to be a separate color channel of RGB where
% Mangular is red, Mtime is blue, Mradial is green. We perform linear interpolation
% on each M individually to extend the probabilities to be within the range of 0 and 255.

\chapter{Main Results}

\chapter{Conclusion}
\label{chap:conclusion}

% \bibliography{../references/main}
% \bibliographystyle{plain}

\end{document}

### 2020-03-10 presentation/slides.tex
initial commit

\documentclass[notes]{beamer}           % print frame + notes
% \documentclass[notes=only]{beamer}    % only notes
% \documentclass{beamer}                % only frames

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{algorithmic,float}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{graphicx}
\usepackage{media9}

% Adds do-while syntax for algorithms
\SetKwRepeat{Do}{do}{while}%

% Case Statements for Proofs
\newcounter{case}[section]
\newenvironment{case}[1][]{\refstepcounter{case}\par\medskip
   \textbf{Case~\thecase. #1} \rmfamily}{\medskip}

\newcounter{subcase}[case]
\newenvironment{subcase}[1][]{\refstepcounter{subcase}\par\medskip
   \textbf{Case~\thecase.\thesubcase. #1} \rmfamily}{\medskip}

% Give names to lemmas
\newtheoremstyle{named}{}{}{\itshape}{}{\bfseries}{.}{.5em}{\thmnote{#3's }#1}
\theoremstyle{named}
\newtheorem*{namedlemma}{Lemma}

% Multi-part proofs
\makeatletter
\newenvironment<>{proofs}[1][\proofname]{%
    \par
    \def\insertproofname{#1\@addpunct{.}}%
    \usebeamertemplate{proof begin}#2}
  {\usebeamertemplate{proof end}}
\makeatother

% Section Title Slides
\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

% Title Stuff
\title{Optimal Pattern Formation for Systems of Mobile Robots}
\author{Jared Coleman}
\institute{California State University, Long Beach}
\date{December 10, 2019}

% Changes multi-slide frame numbering to "Cont."
\setbeamertemplate{frametitle continuation}[from second][(Cont.)]

% Slides
\begin{document}
    \frame{\titlepage}

    \begin{frame}{Introduction}
        \framesubtitle{Motivation}

        Teams of robots often need to form Patterns for exploration, communication, etc.

        The size and orientation of the pattern may not matter, but each robot must conserve energy.
    \end{frame}

    \begin{frame}{Introduction}
        \framesubtitle{The Problem}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{itemize}
                    \item A set of robots with initially arbitrary positions must move to form a pattern.
                    \item Patterns are invariant to translation, rotation, and uniform scaling.
                \end{itemize}
            \end{column}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/intro_formations.png}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}{Model \& Assumptions}
        \begin{itemize}
            \item Robots follow the \textit{Look}, \textit{Compute}, \textit{Move} model
            \begin{itemize}
                \item \textit{Look}: Robots observe the positions of all other robots.
                \item \textit{Compute}: Robots compute their destination in the optimal formation
                \item \textit{Move}: Robots move toward their destination
            \end{itemize}
            \item All robots execute the \textit{LCM} cycle synchronously
            \item Robots move at the same speed
        \end{itemize}
    \end{frame}

    \begin{frame}{Preliminaries}
        \begin{itemize}
            \item Calligraphic notation (e.g. $\mathcal{S}$, $\mathcal{M}$) for sets of sets of points.
            \item Capital letters for sets of points (e.g. $S$, $M$)
            \begin{itemize}
                \item $S^i$ is the $i^{th}$ point in $\mathcal{S}$
            \end{itemize}
            \item lower-case letters for points (e.g. $s$, $m$)
            \begin{itemize}
                \item $s^i_j$ is the $j^{th}$ point in $S^i$
            \end{itemize}
            \item $d(p, q)$ is the euclidian distance between points $p$ and $q$.
            \item $C(p, r)$ is a circle centered at $p$ with radius $r$.
            \item $D(p, r)$ is the closed disk centered at $p$ with radius $r$.
        \end{itemize}
    \end{frame}

    \begin{frame}
        \frametitle{Problem}

        Given two sets of points, $P$ and $S$ that represent the robots' original positions and a desired pattern, respectively, find the minimum maximum distance traveled by any robot necessary to form $S$:

        \begin{align*}
            d^* = \underset{Q \sim S}{min} ~ \underset{0 \leq i < n}{max} \sim d(p_i, q_i)
        \end{align*}

        Let $Q^*$ be the formation that yields the optimal solution $d^*$.
    \end{frame}

    \begin{frame}[allowframebreaks]{Necessary Conditions}
        \framesubtitle{Minimum Rigidness}
        \begin{lemma}
            For any set of points $P$ with optimal solution $d^*$, at least three robots move exactly $d^*$.
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.6\textwidth}
                \begin{proofs}
                    \begin{case}
                        Suppose exactly $1$ robot, $r_0$, moves distance $d^*$ and all others move some distance less than or equal to $d^\prime < d^*$.

                        \bigskip

                        We can achieve a better solution by translating $Q^*$ some arbitrarily small $\epsilon < d^\prime$ so that $d(q_0, p_0) = d^* - \epsilon$ and $d(q_i, p_i) \leq d^\prime + \epsilon < d^*$ for $i \neq 0$.

                        \bigskip

                        This is contradictory to $Q^*$ being optimal.
                    \end{case}
                \end{proofs}
            \end{column}
            \begin{column}{0.4\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/onerobots.pdf}
                \end{center}
            \end{column}
        \end{columns}

        \framebreak

        \begin{columns}
            \begin{column}{0.6\textwidth}
                \begin{proof}[\proofname\ (Cont.)]
                    \begin{case}
                        Now suppose exactly $2$ robots, $r_0$ and $r_1$, move distance $d^*$ and all other robots move some distance less than or equal to $d^\prime < d^*$.

                        \bigskip

                        We can achieve a better solution by scaling and rotating $Q^*$ around $q_0$ some arbitrarily small amount such that $d(q_0, p_0) = d^* - \epsilon$ and $d(q_i, p_i) \leq d^\prime + \epsilon_i < d^*$ for $2 \leq i < n$.

                        \bigskip

                        This reduces to Case 1.
                    \end{case}
                \end{proof}
            \end{column}
            \begin{column}{0.4\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/tworobots.pdf}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \begin{frame}[allowframebreaks]{A Motivating Example}
        Three robots need to form an equilateral triangle. 
        \bigskip

        Suppose we have access to an Oracle that, given a distance $r$, gives us a set of equidistant points $Q$ such that $d(q_i, p_i) = r$, if one exists.

        \framebreak

        A Reasonable solution would be to use a modified binary search:

        \begin{algorithm}[H]
            \algsetup{linenosize=\tiny}
            \SetKwFunction{FApproximateOptimalSolution}{ApproximateOptimalSolution}
            \SetKwProg{Fn}{Function}{:}{}

            \Fn{\FApproximateOptimalSolution{$P$, $\epsilon$}}{
                $low \gets 0$\;
                $high \gets \underset{i,j \in [0, |P|)}{max} d(r_i, r_j)$\;
                \Do{$high-low > 2\epsilon$} {
                    $r \gets \frac{high-low}{2}$\;
                    $Q \gets Oracle(r)$\;
                    \eIf{$Q$ is null}{
                        $low = r$\;
                    }{
                        $high = r$\;
                    }
                } 
                \Return $Q$\;
            }
        \end{algorithm}

        \framebreak

        \begin{itemize}
            \item \texttt{ApproximateOptimalSolution} runs in 
            $\mathcal{O}\left(log\left(\frac{d}{\epsilon}\right)\right)$ 
            where $d$ is the greatest distance between any two robots and $\epsilon$ is the acceptable error for the solution.
            \item \texttt{ApproximateOptimalSolution} is an \textit{approximation} - \textbf{not optimal}.
            \item Each robot is guaranteed to move no more than $\epsilon$ further than in the optimal solution.
            \item \textbf{Spoiler}: We propose a fixed-step, ruler-and-compass method for computing the optimal solution
        \end{itemize}
    \end{frame}

    \section{Three Robots}

    \begin{frame}[allowframebreaks]{Replication}
        \begin{itemize}
            \item A tool we based on pure geometry that resembles a pantograph.
            \item Used to derive later results.
        \end{itemize}

        \framebreak

        \begin{definition}[Trivial Replication]
            The Trivial Replication, $R_{Trivial}(P, q_0, q_1)$, is the set of points similar to $P$ whose first and second point are $q_0$ and $q_1$, respectively.
        \end{definition}

        \framebreak

        \begin{columns}
            \begin{column}{0.7\textwidth}
                \begin{definition}[Replication Machine]
                    The Replication Machine, $R_{Machine}(P, q_0, q_1, r)$, is the infinite set of shapes similar to $P$ whose first point is $q_0$ and second point is on the circumference of the circle centered at $q_1$ with radius $r$.                     
                \end{definition}
            \end{column}
            \begin{column}{0.3\textwidth}
                \begin{center}
                    \includemedia[
                        width=0.75\textwidth,height=0.75\linewidth,
                        activate=pageopen,
                        transparent,
                        addresource=../media/replication_machine.mp4,
                        flashvars={source=../media/replication_machine.mp4&autoPlay=true&loop=true}
                      ]{}{VPlayer.swf}
                \end{center}
            \end{column}
        \end{columns}

        \framebreak

        \begin{columns}
            \begin{column}{0.7\textwidth}
                \begin{definition}[Replication Spanner]
                    The Replication Spanner, $R_{Span}(P, q_0, q_1, r)$, is the infinite set of shapes similar to $P$ whose first and second points are on the circumferences of the circles with radius $r$ centered at $q_0$ and $q_1$, respectively.                 
                \end{definition}
            \end{column}
            \begin{column}{0.3\textwidth}
                \begin{center}
                    \includemedia[
                        width=0.75\textwidth,height=0.75\linewidth,
                        activate=pageopen,
                        transparent,
                        addresource=../media/replication_spanner.mp4,
                        flashvars={source=../media/replication_spanner.mp4&autoPlay=true&loop=true}
                      ]{}{VPlayer.swf}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \note{
        Add picture of pantograph (and a little history)
    }

    \begin{frame}[allowframebreaks]{Replication Machine Circles}

        Given a formation $P$ of size $n$, points $q_0$ and $q_1$, and a distance $r$.

        \begin{lemma}
            The $n$ sets of $i^{th}$ points in the replicated shapes each form a circle centered at $t_i$ with radius 
            $r\frac{d(q_0, t_i)}{d(q_0, q_1)}$.
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}       
                \begin{proofs}
                    Let $T = R_{Trivial}(P, q_0, q_1)$ and $\mathcal{M} = R_{Machine}(P, q_0, q_1, r)$.

                    \bigskip

                    Observe that since $T$ is similar to $M \in \mathcal{M}$, then for some k,
                    $d(m_0, m_1) = k ~ d(t_0, t_1)$ and 
                    $d(m_0, m_i) = k ~ d(t_0, t_i)$.

                    \bigskip

                    Therefore, $d(t_i, m_i) = k ~ d(t_1, m_1)$.

                    \bigskip

                    Since $k ~ d(t_1, m_1)$ is constant, 
                    The set forms a circle with radius $d(t_i, m_i)$
                    centered at $t_i$.
                \end{proofs}        
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/replication_machine.png}
                \end{center}
            \end{column}
        \end{columns}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{proof}[\proofname\ (Cont.)]
                    Now consider the replication $M$ such that $m_1$
                    is colinear with $\overline{q_0q_1}$.

                    \bigskip

                    Observe that $k d(q_0, q_1) = d(q_0, q_1) + r$ and 
                    $k d(q_0, t_i) = d(q_1, t_1) + d(t_i, m_i)$.

                    \bigskip

                    Solving the system of equations results in 
                    $d(t_i, m_i) = r \frac{d(q_0, t_i)}{d(q_0, q_1)}$.
                \end{proof}  
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/replication_machine_radius.png}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \begin{frame}[allowframebreaks]{Replication Spanner Circles}

        Given a formation $P$ of size $n$, points $q_0$ and $q_1$, and a distance $r$.

        \begin{lemma}
            The $n$ sets of $i^{th}$ points in the replicated shapes are contained entirely in a closed disk centered at $t_i$ with radius 
            $r \frac{d(q_0, t_i) + d(q_1, t_i)}{d(q_0, q_1)}$.
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{proof}

                    Let $T = R_{Trivial}(P, q_0, q_1)$, 
                    $\mathcal{M} = R_{Machine}(\{p_1, p_0, p_2, \ldots, p_{n-1}\}, q, p, r)$, and 
                    $\mathcal{S} = R_{Span}(P, p, q, r)$

                    \bigskip

                    Observe the center of the $i^{th}$ circle spanned by $S$
                    is in the $i^{th}$ circle spanned by $\mathcal{M}$. 
                    Therefore, $t_i$ is a center-of-centers.

                    The radius can then be computed:

                    \begin{center}
                       $d(t_i, m_i) + d(m_i, s_i) = r \frac{d(q_0, t_i) + d(q_1, t_i)}{d(q_0, q_1)}$
                    \end{center}
                 \end{proof} 
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/replication_spanner.png}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \section{Forming Triangles}

    \begin{frame}{Construction}
        \framesubtitle{Step 1: Trivial Replications}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                Given the initial positions $P$ and pattern $S$, consider all of the trivial replications of $S$ on $P$, 
                $t^i = T(\{s_{i+1}, s_{i+2}, s_i\}, p_{i+1}, p_{i+2})_i$ 
                for $0 \leq i \leq 2$.
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/construction_1.png}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}{Construction}
        \framesubtitle{Step 2: Replication Spanners}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                Then, for some unknown $r$, consider the replication spanners
                $R_{SPAN}(\{s_{i}, s_{i+1}, s_{i+2}\}, p_i, p_{i+1}, r)$
                for $0 \leq i \leq 2$.

                \bigskip

                Notice that for some $r$ there exists a single point of intersection, $q_i$, between $C(p_i, r)$ and the $i^{th}$ circle generated by $\mathcal{S}$.

            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/construction_2.png}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}{Construction}
        \framesubtitle{Step 3: Compute the Radius}
        Then $r$ can be computed algebraically:

        \begin{align*}
            d(p_i, t_i) - r &= r \frac{d(s_i, s_{i+1}) + d(s_i, s_{i+2})}{d(s_{i+1}, s_{i+2})} \\ 
            d(p_i, t_i) &= r \left( \frac{d(s_i, s_{i+1}) + d(s_i, s_{i+2})}{d(s_{i+1}, s_{i+2})} + 1 \right) \\ 
            d(p_i, t_i) &= r \left( \frac{d(s_i, s_{i+1}) + d(s_i, s_{i+2}) + d(s_{i+1}, s_{i+2})}{d(s_{i+1}, s_{i+2})} \right) \\
            d(p_i, t_i) &= \frac{r}{d(s_{i+1}, s_{i+2})} \\
            r &= d(p_i, t_i) ~ d(s_{i+1}, s_{i+2})
        \end{align*}
    \end{frame}

    \begin{frame}[allowframebreaks]{Solution Validity}
        \begin{lemma}
            The target formation $Q = \{q_0, q_1, q_2\}$ is similar to the desired pattern $S$.
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{proof}
                   Note that given $q_2$, there must exist a unique pair of points 
                   $q^\prime_0 \in C(p_0, r)$ and 
                   $q^\prime_1 \in C(p_1, r)$ such that 
                   $Q^\prime=\{q^\prime_0, q^\prime_1, q_2\} \sim S$.

                   \bigskip

                   Suppose $q^\prime_1 \neq q_1$. This implies 
                   $q^\prime_1 \not\in D \left( p^\prime_1, r \frac{d(q_0, p^\prime_1) + d(q_1, p^\prime_1)} {d(q_0, q_1)} \right)$.

                   \bigskip

                   This is a contradiction.
                \end{proof}

            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/construction_valid.png}
                \end{center}
            \end{column}
        \end{columns}

    \end{frame}

    \note{
        Fact follows from the results of the translation spanner defined above. \\
        A similar argument can be made to prove $q_0 = q^\prime_0$
    }

    \begin{frame}[allowframebreaks]{Solution Optimality}
        \begin{lemma}
            The target formation $Q = \{q_0, q_1, q_2\}$ is the optimal solution for the initial positions $P$.
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{proofs}
                    Let $r^\prime$ be some distance arbitrarily smaller than $r$.
                    Let $\mathcal{S}=R_{Span}(\{s_{i+1}, s_{i+2}, s_i\}, p_{i+1}, p_{i+2}, r^\prime)$.

                    \bigskip

                    \textbf{Recall}: $\mathcal{S}$ is the set of all formations similar to $P$ such that $r_{i+1}$ and $r_{i+2}$ travel distance $r$.
                \end{proofs}

            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/construction_optimal.png}
                \end{center}
            \end{column}
        \end{columns}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{proof}[\proofname\ (Cont.)]
                    Notice, that since $r^\prime < r$, then $C(p_i, r^\prime)$ and the $i^{th}$ circle generated by $\mathcal{S}$ do not intersect.

                    \bigskip

                    Therefore, there exists no $Q^\prime \in \mathcal{S}$ such that $q_i \in C(p_i, r^\prime)$.

                    \bigskip

                    This is a contradiction.
                \end{proof}

            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/construction_optimal.png}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}{Assignment}
        \begin{itemize}
            \item Previous solution assumes a prescribed assignment between robots and their relative position in the pattern.
            \item Must test all possible assignments.
            \item Systems of three robots have $6$ possible assignments.
            \item Later, we introduce an optimization to avoid computing for all $6$ assignments.
        \end{itemize}    
    \end{frame}

    \begin{frame}{Observations}
        \framesubtitle{Focal Point}

        \begin{theorem}
            Given a set of three robots arbitrarily deployed on the plane at positions $P$, a pattern $S$, and the optimal solution, $Q^*$ there exists a single point that every robot moves either toward or away from.
        \end{theorem}

    \end{frame}

    \begin{frame}[allowframebreaks]{Observations}
        \framesubtitle{Center-of-Mass Invariant}

        \begin{lemma}
            When $S$ forms an equilateral triangle, the center of mass of the original formation is the same as the destination formation.
            Furthermore, if every robot moves at the same speed, then the center of mass of all three robots never changes as they move toward their final destination in the formation. 
        \end{lemma}

        \framebreak

        \begin{columns}
            \begin{column}{0.5\textwidth}      
                \begin{proofs}
                    Note the center of mass is: 
                    $c = \left( \frac{r^x_0 + r^x_1 + r^x_2}{3}, \frac{r^y_0 + r^y_1 + r^y_2}{3}\right)$

                    \bigskip

                    Suppose $r_0$ and $r_1$ move towards the origin and $r_1$ moves away from the origin with a speed of $1$. 
                \end{proofs}
            \end{column}
            \begin{column}{0.5\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=\textwidth]{../media/center_of_mass.png}
                \end{center}
            \end{column}
        \end{columns}

        \framebreak

        \begin{columns}
            \begin{column}{0.6\textwidth}     
                \begin{proof}[\proofname\ (Cont.)]

                    Then, after a single time step:

                    \begin{align*}
                        c^\prime &= \left( \frac{r^x_0 + (r^x_1 + cos \frac{\pi}{6}) + (r^x_2 - cos \frac{\pi}{6})}{3},\right. \\
                        &\quad \quad \left. \frac{(r^y_0 + 1) + (r^y_1 - sin \frac{\pi}{6}) + (r^y_2 - sin \frac{\pi}{6})}{3}\right) \\
                        &= \left( \frac{r^x_0 + r^x_1 + r^x_2}{3}, \frac{r^y_0 + r^y_1 + r^y_2}{3}\right) = c
                    \end{align*}
                \end{proof}
            \end{column}
            \begin{column}{0.4\textwidth}  %%<--- here
                \begin{center}
                    \includegraphics[width=0.7\textwidth]{../media/center_of_mass.png}
                \end{center}
            \end{column}
        \end{columns}
    \end{frame}

    \begin{frame}[allowframebreaks]{Observations}
        \framesubtitle{Invariance to Obliviousness}

        \begin{lemma}
            If all robots move at the same speed, any optimal solution is invariant with respect to time.
        \end{lemma}

        \framebreak

        \begin{proofs}
            Let $t^*$ be the time that every robot reaches its destination, or $f_i(t^*) = p_i$.

            Let $Q^{t}$ bethe optimal solution computed at time $t$.

            Suppose at time $t < t^*$, the robot has moved some distance 
            $r < d(p_i, q_i)$.

            Let $\hat{r}$ be the distance the robot has left to travel, should it continue towards $q_i$, or $\hat{r} = d(p_i, q_i) - r$.
        \end{proofs}

        \framebreak

        \begin{proof}[\proofname\ (Cont.)]
            First, note that 
            $d(p_i, q^{t}_i) \geq d(p_i, q_i)$, 
            or else $Q^*$ would not be optimal for the initial configuration.

            Also, notice that
            $d(f_i(t), q^{t}_i) \leq d(f_i(t), q_i^*)$, or else
            $Q*$ is a better solution than $Q^{t}$, which is a contradiction to the assumption that $P^{t}$ is optimal.

            The only point where both of these conditions are satisfied is 
            $P^{t} = P^*$
        \end{proof}

        \framebreak

        \begin{center}
            \includemedia[
                width=0.7\textwidth,height=0.7\linewidth,
                activate=pageopen,
                transparent,
                addresource=../media/solution_set.mp4,
                flashvars={source=../media/solution_set.mp4&autoPlay=true&loop=true}
                ]{}{VPlayer.swf}
        \end{center}
    \end{frame}

    \begin{frame}[allowframebreaks]{Observations}
        \framesubtitle{Center Invariant}

        \begin{theorem}
            The center-of-mass of optimal solution is the incenter of circumcenters of the trivial replications used to compute it.
        \end{theorem} 
    \end{frame}

    \section{The General Case}

    \begin{frame}{The General Case}
        \framesubtitle{A Partial Solution}

        \begin{itemize}
            \item We know at least $3$ robots must move distance $d^*$
            \item There are $\binom{n}{3} \times \binom{n}{3} = \mathcal{O}(n^6)$ possible assignments where exactly $3$ robots move $d^*$.
            \item Thus, in the case where exactly $3$ robots move distance $d^*$, a solution can be found in $\mathcal{O}(n^8)$.
            \begin{itemize}
                \item Using an $\mathcal{O}(n^2)$ min-max maximum cardinality matching for each possible assignment.
            \end{itemize}
        \end{itemize}

    \end{frame}

\end{document}

### 2020-03-10 code/collinear.py

@@ -1,0 +1,15 @@
+import numpy as np
+from shapes import random_shape
+
+def random_example():
+    points = np.random.random_sample((3, 2))
+    speeds = np.random.random_sample((3, 2))
+
+    speeds = np.linalg.norm(speeds, axis=0) / 1e-3
+    
+    positions = np.arange(1, 1001)[:,np.newaxis] * speeds 
+
+    print(positions)
+
+if __name__ == '__main__':
+    random_example()

### 2020-03-10 code/formation.py

@@ -1,0 +1,202 @@
+import numpy as np 
+import math
+
+import matplotlib.pyplot as plt
+
+from functools import partial 
+from itertools import permutations, combinations
+from typing import Callable, Dict, Tuple, List
+
+from plotters import *
+
+def get_transform(v1: np.array, v2: np.array) -> float:
+    norm_1 = np.linalg.norm(v1)
+    norm_2 = np.linalg.norm(v2)
+    angle = math.acos(np.dot(v1, v2) / (norm_1 * norm_2))
+    scale = norm_2 / norm_1
+
+    if np.linalg.det([v1.astype(np.float32), v2.astype(np.float32)]) > 0:
+        angle = -angle
+
+    return angle, scale
+
+def replication(formation: np.array, 
+                q_0: np.array, q_1: np.array, 
+                i: int=0, j: int=1) -> np.array:
+    """Returns the replicated shape.
+
+    This returns the shape P similar to formation such that
+    P's ith and jth points are fixed to q_0 and q_1, respectively.
+
+    Args:
+        formation (np.array): formation to replicate
+        q_0: Point to fix ith point in target formation to
+        q_1: Point to fix jth point in target formation to
+        i: index of point in target formation to fix to q_0
+        j: index of point in target formation to fix to q_1
+
+    Returns:
+        np.array: The replicated shape
+    """
+    formation = np.copy(formation)
+    q_0 = np.copy(q_0)
+    q_1 = np.copy(q_1)
+
+    assert(len(formation.shape) == 2)   # two-dimensional
+    assert(formation.shape[1] == 2)     # collection of two-dimensional points
+    assert(q_0.shape == (2,))           # two-dimensional point
+    assert(q_1.shape == (2,))           # two-dimensional point
+
+    formation -= (formation[i] - q_0)   # shift to i to q_0
+
+    offset = np.copy(formation[i])
+    formation -= offset     # temp shift i to origin
+    q_1 -= offset
+
+    theta, scale = get_transform(formation[j], q_1)
+
+    c = math.cos(-theta)
+    s = math.sin(-theta)
+
+    formation = np.array(
+        [
+            formation[:,0] * c - formation[:,1] * s,
+            formation[:,0] * s + formation[:,1] * c
+        ]
+    ).T * scale + offset
+
+    return formation
+
+def replication_machine_circles(formation: np.array,  
[…]

### 2020-03-10 code/gabriel.py

@@ -1,0 +1,75 @@
+import numpy as np
+import math
+from itertools import product, chain, combinations, permutations, starmap
+from functools import partial
+
+def dist(p1, p2):
+    return math.sqrt((p1[0] - p2[0])**2 + (p1[1] - p2[1])**2)
+
+def mid(p1, p2):
+    return (p1[0] + p2[0]) / 2, (p1[1] + p2[1]) / 2
+
+def get_gabriel_graph(points: np.array, points2: np.array = None):
+    if points2 is None:
+        points2 = points
+    aux = []
+    for p1, p2 in product(points, points2):
+        if p1 == p2:
+            continue
+        m = mid(p1, p2)
+        aux.append((m, dist(m, p1), p1, p2))
+
+    edges = []
+    for m, rad, p1, p2 in aux:
+        if (True in map(lambda p: dist(m, p) < rad, points) and
+            True in map(lambda p: dist(m, p) < rad, points2)):
+            continue
+        edges.append((p1, p2))
+
+    return edges
+
+def brute_match(points, points2):
+    edges = []
+    r = None
+    for _points2 in permutations(points2):
+        _r = max(starmap(dist, zip(points, _points2)))
+        if r is None or _r < r:
+            r = _r
+            edges = list(zip(points, _points2))
+    return edges
+
+import networkx as nx
+import matplotlib.pyplot as plt
+from matplotlib import collections  as mc
+
+if __name__ == '__main__':
+    points = list(map(tuple, np.random.rand(5, 2)))
+    points2 = list(map(tuple, np.random.rand(5, 2)))
+    edges = set(get_gabriel_graph(points, points2))
+
+    print('Checking Brutely')
+    brute_edges = set(brute_match(points, points2))
+    if brute_edges.issubset(edges):
+        print('TRUTH')
+    else:
+        print('FALSE')
+
+    fig: plt.Figure
+    ax: plt.Axes
+    fig, ax = plt.subplots()
+    ax.scatter(*zip(*points), color='b')
+    ax.scatter(*zip(*points2), color='y')
+
+    lc = mc.LineCollection(edges)
+    ax.add_collection(lc)
+    
+    lc_brute = mc.LineCollection(brute_edges, color='g')
+    ax.add_collection(lc_brute)
+    
+    lc_absent = mc.LineCollection(brute_edges - edges, color='r')
+    ax.add_collection(lc_absent)
+
+    ax.autoscale()
+    ax.margins(0.1)
+    plt.show()
+

### 2020-03-10 code/geometry.py

@@ -1,0 +1,39 @@
+import numpy as np
+from math import sqrt
+
+def circle_intersection(circle1, circle2):
+        '''
+        @summary: calculates intersection points of two circles
+        @param circle1: tuple(x,y,radius)
+        @param circle2: tuple(x,y,radius)
+        @result: tuple of intersection points (which are (x,y) tuple)
+        '''
+        # return self.circle_intersection_sympy(circle1,circle2)
+        x1,y1,r1 = circle1
+        x2,y2,r2 = circle2
+        # http://stackoverflow.com/a/3349134/798588
+        dx,dy = x2-x1,y2-y1
+        d = sqrt(dx*dx+dy*dy)
+        if d > r1+r2:
+            print("#1")
+            return None # no solutions, the circles are separate
+        if d < abs(r1-r2):
+            print("#2")
+            return None # no solutions because one circle is contained within the other
+        if d == 0 and r1 == r2:
+            print("#3")
+            return None # circles are coincident and there are an infinite number of solutions
+
+        a = (r1*r1-r2*r2+d*d)/(2*d)
+        h = sqrt(r1*r1-a*a)
+        xm = x1 + a*dx/d
+        ym = y1 + a*dy/d
+        xs1 = xm + h*dy/d
+        xs2 = xm - h*dy/d
+        ys1 = ym - h*dx/d
+        ys2 = ym + h*dx/d
+
+        return np.array([
+            [xs1, ys1],
+            [xs2, ys2]
+        ])

### 2020-03-10 code/helpers.py

@@ -1,0 +1,19 @@
+import numpy as np
+
+def get_intersect(a1, a2, b1, b2):
+    """ 
+    Returns the point of intersection of the lines passing through a2,a1 and b2,b1.
+    a1: [x, y] a point on the first line
+    a2: [x, y] another point on the first line
+    b1: [x, y] a point on the second line
+    b2: [x, y] another point on the second line
+    """
+    s = np.vstack([a1, a2, b1, b2])        # s for stacked
+    h = np.hstack((s, np.ones((4, 1))))    # h for homogeneous
+    l1 = np.cross(h[0], h[1])              # get first line
+    l2 = np.cross(h[2], h[3])              # get second line
+    x, y, z = np.cross(l1, l2)             # point of intersection
+    if z == 0:                             # lines are parallel
+        return np.array([float('inf'), float('inf')])
+    return np.array([x/z, y/z])
+

### 2020-03-10 code/optimization.py

@@ -1,0 +1,169 @@
+from cvxopt import matrix, solvers
+from math import sqrt 
+import numpy as np
+import matplotlib.pyplot as plt
+from itertools import permutations, combinations, chain
+from plotters import get_colors
+
+from typing import List
+
+from formation import triangle_solution, triangle_solution_no_perm, replication
+from shapes import random_shape, regular_shape, Triangle
+solvers.options['show_progress'] = False
+
+def print_arrays(**kwargs):
+    name: str
+    value: np.array
+    for name, value in kwargs.items():
+        print(f'{name} {value.shape}')
+        print(value)
+        print()
+
+def format_eqs(lefts: List[List[float]], right: List[float]):
+    *xs, cs = zip(*([right] + lefts))    
+    mat = []
+    for elems in xs:
+        mat.append([float(-elem) for elem in elems])
+
+    return matrix(mat), matrix(cs)
+
+import math
+def convex_solution_no_perm(positions: np.array, formation: np.array, verbose: bool = False) -> np.array:
+    positions = np.copy(positions)
+    formation = np.copy(formation) - formation[0] # adjust formation to origin
+    angle = -math.atan2(formation[1][1], formation[1][0])
+    xs, ys = formation[:,0].copy(), formation[:,1].copy()
+    formation[:,0] = math.cos(angle) * xs - math.sin(angle) * ys
+    formation[:,1] = math.sin(angle) * xs + math.cos(angle) * ys
+
+    # assert((formation[0] == np.array([0.0, 0.0])).all())
+    assert(positions.shape == formation.shape)
+
+    s2norm = np.linalg.norm(formation[1])
+    n = len(formation)
+
+    # | q_i.x - p_i.x |    
+    # |               |     <= r
+    # | q_i.y - p_i.y |_2 
+    # 
+    # such that 
+    # Aq = 0    <---    this says the shape is similar
+
+    Gs = []
+    hs = []
+
+    eqs = np.append(np.eye(2*n), np.zeros((2*n, 2)), axis=1)
+    eqs[:,-1][::2] = -positions[:,0]
+    eqs[:,-1][1::2] = -positions[:,1]
+
+    eq_r = np.zeros(2*n + 2)
+    eq_r[-2] = 1.0
+    for i in range(n):
+        G, h = format_eqs(
+            [eqs[2*i].tolist(), eqs[2*i+1].tolist()],
[…]

### 2020-03-10 code/plotters.py

@@ -1,0 +1,33 @@
+
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+import numpy as np 
+
+from typing import Iterable
+
+def get_colors(num: int) -> Iterable[str]:
+    return cm.rainbow(np.linspace(0, 1, num))
+
+def plot_shape(ax: plt.Axes, shape: np.array, color: str = 'black', marker: str = 'o') -> None:
+    ax.scatter(x=shape[:,0], y=shape[:,1], color=color, marker=marker)
+
+def plot_shapes(ax: plt.Axes, *shapes: np.array, marker: str = 'o') -> None:
+    colors = get_colors(len(shapes))
+    for shape, c in zip(shapes, colors):
+        plot_shape(ax, shape, color=c, marker=marker)
+
+def plot_polygon(ax: plt.Axes, polygon: np.array, color: str = 'black') -> None:
+    ax.add_patch(plt.Polygon(polygon, color=color, fill=False))
+
+def plot_polygons(ax: plt.Axes, *polygons: np.array) -> None:
+    colors = get_colors(len(polygons))
+    for polygon, color in zip(polygons, colors):
+        plot_polygon(ax, polygon, color=color)
+
+def plot_circle(ax: plt.Axes, point: np.array, radius: float, color: str = 'black') -> None:
+    ax.add_patch(plt.Circle(point, radius, color=color, fill=False))
+
+def plot_circles(ax: plt.Axes, points: np.array, radiuses: np.array) -> None:
+    colors = get_colors(len(points))
+    for point, radius, color in zip(points, radiuses, colors):
+        plot_circle(ax, point, radius, color=color)

### 2020-03-10 code/shapes.py

@@ -1,0 +1,39 @@
+import numpy as np 
+import math 
+
+def random_shape(num: int, low: float = -1.0, high: float = 1.0) -> np.ndarray:
+    return np.random.random_sample((num, 2)).astype(np.longdouble) * (high - low) + low
+
+def regular_shape(num: int) -> np.ndarray:
+    return np.array(
+        [
+            [math.cos(2 * math.pi * i / num), math.sin(2 * math.pi * i / num)]
+            for i in range(num)
+        ]
+    ).astype(np.longdouble)
+
+class Triangle:
+    EQUILATERAL = np.array([
+        [0.0, 0.0],
+        [1.0, 0.0],
+        [0.5, math.sqrt(3) / 2]
+    ]).astype(np.longdouble)
+    
+    @staticmethod
+    def right(angle: float, deg: float = False) -> np.ndarray:
+        if deg:
+            angle = angle * math.pi / 180
+        return np.array([
+            [0.0, 0.0],
+            [1.0, 0.0], 
+            [1.0, math.tan(angle)]
+        ]).astype(np.longdouble)
+
+    @staticmethod
+    def isosoles(height: float) -> np.ndarray:
+        return np.array([
+            [-0.5, 0.0],
+            [0.0, height],
+            [0.5, 0.0],
+        ]).astype(np.longdouble)
+

### 2020-03-10 code/solution.py

@@ -1,0 +1,129 @@
+from formation import triangle_solution_no_perm, replication
+from itertools import product, permutations, combinations
+
+import numpy as np
+
+def is_similar(shp, _shp):
+    sides = np.linalg.norm(shp - np.roll(shp, 1, axis=0), axis=1)
+    sides =  np.sort(sides) / np.sum(sides)
+    _sides = np.linalg.norm(_shp - np.roll(_shp, 1, axis=0), axis=1)
+    _sides = np.sort(_sides) / np.sum(_sides)
+    return np.isclose(sides, _sides).all()
+
+def solution(positions: np.array, formation: np.array) -> np.array:
+    combos = lambda x: map(np.array, combinations(range(x), 3))
+    n = len(positions)
+    target, radius = None, None
+    for i, j in product(combos(n), combos(n)):
+        pos = positions[i]
+        for perm in map(np.array, permutations(range(3))):
+            tri_target, tri_radius = triangle_solution_no_perm(pos, formation[j[perm]])
+                
+            _target = replication(formation, tri_target[0], tri_target[1], j[perm][0], j[perm][1])
+
+            tmp_radius = None
+            for _tar in map(np.array, permutations(_target)):
+                _radiuses = np.linalg.norm(positions - _tar, axis=1)
+                _radius = np.max(_radiuses)
+                if radius is None or _radius < radius:
+                    target, radius = _tar, _radius
+                if tmp_radius is None or _radius < tmp_radius:
+                    tmp_radius = _radius
+    return target
+
+from shapes import random_shape
+from optimization import convex_solution, convex_solution_no_perm
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+# if __name__ == '__main__':
+#     n, m = 50, 20
+#     data = np.zeros((n, m))
+#     for i in range(3, n):
+#         total = 0
+#         for j in range(m):
+#             positions, formation = random_shape(i), random_shape(i)
+#             target, radius = convex_solution(positions, formation)
+#             radiuses = np.linalg.norm(target - positions, axis=1)
+#             num = len(radiuses[np.isclose(radiuses, radius)])
+#             print(f'{i}\t{j}\t- {num}')
[…]

### 2020-03-10 code/tests.py

@@ -1,0 +1,436 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import math
+from itertools import combinations, product
+from matplotlib.animation import FuncAnimation
+from matplotlib.widgets import Slider, Button, RadioButtons
+from matplotlib.collections import PatchCollection
+
+from multiprocessing import Pool
+from functools import partial
+import signal
+
+from formation import (
+    replication,
+    replication_machine,
+    replication_machine_circles,
+    replication_spanner,
+    replication_spanner_circles,
+    triangle_solution,
+    triangle_solution_no_perm,
+    approximation,
+    get_sides
+)
+from geometry import circle_intersection
+from plotters import *
+from shapes import Triangle, random_shape, regular_shape
+from typing import Tuple, List, Optional
+
+# https://stackoverflow.com/questions/3252194/numpy-and-line-intersections/3252222
+def get_intersect(a1, a2, b1, b2):
+    """
+    Returns the point of intersection of the lines passing through a2,a1 and b2,b1.
+    a1: [x, y] a point on the first line
+    a2: [x, y] another point on the first line
+    b1: [x, y] a point on the second line
+    b2: [x, y] another point on the second line
+    """
+    s = np.vstack([a1, a2, b1, b2])        # s for stacked
+    h = np.hstack((s, np.ones((4, 1))))    # h for homogeneous
+    l1 = np.cross(h[0], h[1])              # get first line
+    l2 = np.cross(h[2], h[3])              # get second line
+    x, y, z = np.cross(l1, l2)             # point of intersection
+    if z == 0:                             # lines are parallel
+        return np.array([float('inf'), float('inf')])
+    return np.array([x / z, y / z])
+
+
+def test_same_solution(positions: Optional[np.array] = None,
+                       formation: Optional[np.array] = None,
+                       speed: float = 0.002):
+    if positions is None:
+        positions = random_shape(3)
+        print(positions)
+    if formation is None:
+        formation = regular_shape(3)
+
+    fig: plt.Figure
+    ax: plt.Axes
+    fig, ax = plt.subplots()
+    plt.subplots_adjust(bottom=0.25)
+    ax.axis('equal')
+    
+    axslider = plt.axes([0.25, 0.1, 0.65, 0.03], facecolor='lightgoldenrodyellow')
[…]

### 2020-03-11 paper/__latexindent_temp.tex
enter

\documentclass{turabian-thesis}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{pslatex}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{paralist}
\usepackage[utf8]{inputenc}
\usepackage{multicol} 
\usepackage[hidelinks]{hyperref}
\usepackage{nameref}

\DeclareUnicodeCharacter{2215}{\textminus}
\usepackage{algorithm}
\usepackage{algpseudocode}
\algdef{SE}[DOWHILE]{Do}{DoWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{claim}{Claim}
\newtheorem{definition}{Definition}
\newtheorem{problem}{Problem}

\newcounter{case}[section]
\newenvironment{case}[1][]{\refstepcounter{case}\par\medskip
   \textbf{Case~\thecase. #1} \rmfamily}{\medskip}

\newcounter{subcase}[case]
\newenvironment{subcase}[1][]{\refstepcounter{subcase}\par\medskip
   \textbf{Case~\thecase.\thesubcase. #1} \rmfamily}{\medskip}

\begin{document}

\frontmatter
\thispagestyle{empty}

\begin{center}

   \textbf{hybrid model}

   \vspace*{\baselineskip}

   A THESIS \\
   Presented to the Department of Computer Science and Computer Engineering \\
   California State University, Long Beach

   \vspace*{\baselineskip}

   In Partial Fulfillment \\
   of the Requirements for the Degree \\
   Master of Science in Computer Science \\
   Option in Computer Science

   \vspace*{\baselineskip}

   Committee Members: \\

   \vspace*{\baselineskip}

   By Ruben Rosales \\
   May 2020
\end{center}

\pagebreak

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\tableofcontents
\listofillustrations

\mainmatter

\chapter{Introduction}
[explain what black box and white box is]
In this thesis, we study interpretable machine learning as applied to complex-valued time series. Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.

[add more reasons] 
[add how most time series are classified]
These methods, however, fall short of allowing users to visualize patterns within their dataset and ().

To address this issue, we propose an interpretable hybrid  model that can be extended to any time series dataset. Our model consists of two models. The first is an ensemble method of classifiers that functions as a black box method. The second is a white-box model that encodes time series as images and outputs all images that have been correctly classified.

We study the adaption of deep learning models to complex-valued radio signals. We compare the deep learning models such as CNN’s have become ever-increasingly powerful and widely used but their complex nature makes it difficult for users to understand what is going on so as black-box models get more popular the need for an interpretable accomplice is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand. 
The growth of mobile devices and demand for wireless data has created a need for  High quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.  however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments.
This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically

\section{The need for interpretable models in management problems}

Building upon successful strategies from image and voice recognition domains in machine learning, we demonstrate an approach in the radio using Convolutional Neural Networks (CNNs) and Deep Neural Networks (DNNs) which offers flexibility to learn features across a wide range of tasks and demonstrates improved classification accuracy against current day approaches.  Existing methods, however, fall short of allowing users to visualize patterns within their dataset. To address this issue, we propose a hybrid interpretable model that can be extended to any time series dataset. Our model consists of two models. The first is an ensemble method of classifiers that functions as our black box method. The second is a white-box model that encodes time series as images for each signal and outputs all images that have been correctly classified.

\section{Contributions}
\section{Data}
Our dataset contains two different types of signals, raw and discretized. The raw signal consists of time and amplitude with 31,000 data points while discrete data can vary from 22 to 44. In industry, they do not have access to raw data because it is difficult to track that information so we focus on discrete data for this thesis.

The raw signal has the same time intervals across the entire dataset but it may have different start and end times.

[ What is raw signal and discrete, compare them, ]

\chapter{Preprocessing}
We compared this method to other features extracted from Fourier-transform, fast Fourier transform, sliding window method, and polar coordinates taken directly from the signals amplitude/phase value.

Using any of those methods reduced our models accuracy as well as increased training time. For fourier transform, we attempted to perform ft over the entire signal and use amplitude and time as a feature accuracy dropped below 90%. We believe this is because FT only takes into account the frequency domain and ignores time. FFT and sliding window method gave us better performance averaging around 90% accuracy. We also attempted to take the average of each window and tried different window sizes as well as window overlapping but did not see a significant improvement in loss or accuracy.

We normalize all data using L2 norm and format our data by each subsequent wavelet decomposition. For instance, if we do N levels of decomposition the Nth item would be at the end of the array.

\section{What is it?}
[what is feature extraction]
Data pre-processing is concerned with the analysis and manipulation of the collected spectrum data with the aim to arrive at potentially good wireless data representations. The raw samples organized into data vectors rk in the previous block are pipelined as input for signal processing (SP) tools that analyze, process and transform the
data to arrive at simple data representations such as frequency,
amplitude, phase and spectrum, or more complex features
xk such as e.g. cyclostationary features. In addition, feature
learning such as deep learning may be utilized to automatically
extract more low-level and high-level features. In many ML
applications the choice of features is just as important, if not
more important than the choice of the ML algorithm.

\section{Transformations}
Transformation 1 : STFT/ FT
[what is it]

[fft formula]
The Fourier Transform takes a time-based pattern, measures every possible cycle, and returns the overall "cycle recipe" (the amplitude, offset, and rotation speed for every cycle that was found).

We will begin the analysis of the Fourier Transform on this page. This section gives an introduction to the Fourier Transform and then we take a look at the fundamental properties of the Fourier Transform.

The Fourier Transform is a magical mathematical tool. The Fourier Transform decomposes any function into a sum of sinusoidal basis functions. Each of these basis functions is a complex exponential of a different frequency. The Fourier Transform therefore gives us a unique way of viewing any function - as the sum of simple sinusoids.

The Fourier Series showed us how to rewrite any periodic function into a sum of sinusoids. The Fourier Transform is the extension of this idea to non-periodic functions.

While the the Fourier Transform is a beautiful mathematical tool, its widespread popularity is due to its practical application in virtually every field of science and engineering. It's hard to understand why the Fourier Transform is so important. But I can assure you it enables the solution to difficult problems be made simpler (and also makes previously unsolved problems solvable). In addition, the Fourier Transform gives us a new method of viewing the world, which is fantastic for giving a more intuitive feel for our universe.

[stft]
The Short-time Fourier transform (STFT), is a Fourier-related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.[1] In practice, the procedure for computing STFTs is to divide a longer time signal into shorter segments of equal length and then compute the Fourier transform separately on each shorter segment. This reveals the Fourier spectrum on each shorter segment. One then usually plots the changing spectra as a function of time, known as a spectrogram or waterfall plot.

[why it performed poorly]
The problem with STFT is the fact whose roots go back to what is known as the Heisenberg
Uncertainty Principle. This principle originally applied to the momentum and location of
moving particles, can be applied to time-frequency information of a signal. Simply, this principle
states that one cannot know the exact time-frequency representation of a signal, i.e., one cannot
know what spectral components exist at what instances of times. What one can know are the
time intervals in which certain bands of frequencies exist, which is a resolution problem.

The problem with the STFT has something to do with the width of the window function that is
used. To be technically correct, this width of the window function is known as the support of
the window. If the window function is narrow, then it is known as compactly supported. This
terminology is more often used in the wavelet world, as we will see later.

The problem is that when you are studying a real signal it would be useful to know what the ‘instantaneous frequency’ of the signal is. The instantaneous frequency is the exact frequency of a signal at an exact moment in time. For instance if I was listening to a music track I would like to be able to say ‘at 1 min 59.0423 seconds into the music track the sound is 1563.2 Hz’. Unfortunately the Fourier Transform cannot do this because there exists a minimum amount of uncertainty between the frequency and time domains, like Heisenberg has a minimum amount of uncertainty between the speed and position of a particle (the area of the box). You can know the moment in time you want to find the frequency for (like the blue box), but because there is a minimum uncertainty, the box of has to stretch out across frequency, meaning that you are unsure of the frequency at that moment in time.

The best you can do with a Fourier Transform is to sample a range of time (for instance, the time signal between 1 min 58 sec and 1 min 59 sec in a song) and find a range of frequencies that were played over that amount of time, as represented by the black box. An example of this can be seen by looking at the third picture in the post again. There is a signal over a range of time (e.g. someone saying ‘hello’) and the frequency graph is the range of frequencies recorded over that time.

Transformation 2: Abs/angle
The representation of a complex number as a sum of a real and imaginary number, z = x + iy, is called its Cartesian representation.
We attempt to use the real value of the absolute cartesian representation so we have two features, angle and absolute value,

[purpose and motivation? Behind making it ang/abs]

[definition plus method]

To overcome this resolution problem a Wavelet Transform is used to deconstruct the signal into a load of wavelets being added together. Wavelets are useful because they are limited in time and frequency.  Instead of a wavelet lasting forever and having no limit in time, it dies quickly, like the example of different wavelets below shows:

These waves are limited in time, whereas sin() and cos() are not because they continue forever. When a signal is deconstructed into wavelets rather than sin() and cos() it is called a Wavelet Transform. The graph that can be analysed after the transform is in the wavelet domain, rather than the frequency domain.
This time limiting quality about wavelets is useful to Engineers as it provides more resolution in the time domain. Instead of modelling with an infinite wave, it is possible to model with a finite wave which you can ‘slide’ along to time domain. The ability to slide the signal is what gives Engineers a more accurate representation of the signal and therefore a better resolution in time.

Transformation 2: Sliding window 
We propose a method similar to STFT but in an effort to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.

Transformation 3: Wavelets
% The main advantages of wavelets is that they have a varying window size, being wide for slow frequencies and narrow for the fast ones, thus leading to an optimal time–frequency resolution in all the frequency ranges. Furthermore, owing to the fact that windows are adapted to the transients of each scale, wavelets lack the requirement of stationary [14], [15]. The use of wavelets for PQ studies is most appropriate because most disturbances on a power supply system are non-stationary transients [10], [16]. It can be defined as follows:
% (1)
% where i and j are integers, the functions i,j(x) are the wavelet expansion functions and the two-parameter expansion coefficients ai,j are called the discrete wavelet transform (DWT) coefficients of f(x). The coefficients are given by
% (2)
% The wavelet basis functions can be computed from a function i,j(x) called the generating or mother wavelet through translation and scaling (dilation) parameters

% parameter. Mother wavelet function is not unique, but it must satisfy a small set of conditions. One of them is multiresolution condition and related to the two-scale difference equation

% (4)
% where (x) is scaling function and h(k) must satisfy several conditions to make basis wavelet functions unique, orthonormal and have a certain degree of regularity. The mother wavelet is related to the scaling function as follows:
% (5)
% % where g(k) = (−1) kh(1 − k). At this point, if valid h(x) is available, one can obtain g(x). Note that h and g can be viewed as filter coefficients of half band low-pass and high-pass filters, respectively. J-level wavelet decomposition can be computed with 6 as follows:

% (6)
% where coefficient a0,k and coefficients aj+1,n and dj+1,n at scale j + 1 are given and they can be obtained if coefficient at scale j is available
% (7)
% where a set of wavelet approximation coefficients, aj+1,n, and detailed coefficients dj+1,n at scale j + 1 are given.

Db4 wavelet http://wavelets.pybytes.com/wavelet/db4/

Model 1:
Model 2: 
Model 3:

[Shape of data comparison]

In the preprocessing stage we tried numerous combinations of features but found the best performing features to come from wavelet decomposition.

Wavelets

We implement the discrete wavelet transform as a filter-bank which means it’s composed of cascading high-pass and low-pass filters. This gives us the advantage of splitting a signal into several frequency sub-bands. Wavelet decompositions give us the advantage of gaining features that take into account frequency and time domains.

[explain why they performed better]

[alternatives]

\section{Models}
[introduce models definitions]

This section gives an overview of the theoretical background of concepts discussed in this work.

Machine learning algorithms can simply be seen as learning a target function (f) that maps input data (X) to an output (Y). There are several techniques to make this work but we focus on nonparametric algorithms, including Convolutional Neural Networks (CNN) and Long Short Term Memory networks (LSTM). Nonparametric algorithms are algorithms that attempt to make minimal assumptions about the form of the function so they can learn any form from data provided to it. An example would be a neural network because it has no prior knowledge about what it’s classifying and attempts to generalize any new data points that it has not seen before.
[cnn]

A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis. It consists of an input and output layer as well as hidden layers. A convolution is an operation between a vector of weights w against an input x. It consists of taking the dot product between m and x were in steps of a filter size defined by n.

Convolutional neural networks perform feature learning via non-linear transformations implemented as a series of nested layers. The input data is a multidimensional data array, called tensor, that is presented at the visible layer. This is typically a grid-like topological structure, e.g. time-series data, which can be seen as a 1D grid taking samples at regular time intervals, pixels in images with a 2D layout, a 3D structure of videos, etc. Then a series of hidden layers extract several abstract features. Those layers are “hidden” because their values are not given. Instead, the deep learning model must determine which data representations are useful for explaining the relationships in the observed data. Each layer consists of several kernels that perform a convolution over the input; therefore, they are also referred to as convolutional layers. Kernels are feature detectors, that convolve over the input and produce a transformed version of the data at the output. Those are banks of finite impulse response filters as seen in signal processing, just learned on a hierarchy of layers. The filters are usually multidimensional arrays of parameters that are learnt by the learning algorithm [24] through a training process called backpropagation.

[lstm]
Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture[1] used in the field of deep learning. Unlike standard feedforward neural networks, LSTM has feedback connections. It can not only process single data points (such as images), but also entire sequences of data (such as speech or video). For example, LSTM is applicable to tasks such as unsegmented, connected handwriting recognition[2], speech recognition[3][4] and anomaly detection in network traffic or IDS's (intrusion detection systems).

A common LSTM unit is composed of a cell, an input gate, an output gate and a forget gate. The cell remembers values over arbitrary time intervals and the three gates regulate the flow of information into and out of the cell.

LSTM networks are well-suited to classifying, processing and making predictions based on time series data, since there can be lags of unknown duration between important events in a time series. LSTMs were developed to deal with the exploding and vanishing gradient problems that can be encountered when training traditional RNNs. Relative insensitivity to gap length is an advantage of LSTM over RNNs, hidden Markov models and other sequence learning methods in numerous applications.

[multimodal]
Due to the superior performance and computationally tractable representation capability (in vector spaces) in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks [38, 39, 44]. Typically, domain-specific neural networks are used on different modalities to generate their representations and the individual representations are merged or aggregated. Finally, the prediction is made on top of aggregated representation usually with another a neural network to capture the interactions between modalities and learn complex function mapping between input and output. Addition (or average) and concatenation are two common aggregation methods, i.e.,
The network structure is illustrated in Figure 1(b) . The arrows are function mappings or computing operations. The dotted boxes are representations of single and combined modality features. We call them additive combinations because their critical step is to add modality hidden vectors (although often in a nonlinear way).

[explain architectures]
CNN
Our CNN has an architecture consisting of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X). We found no improvements in performance when we attempted to make it deeper by adding additional convolutional layers as well as skip connections.

LSTM
We wanted to keep our LSTM network as small as possible, for X purposes, so we went with two LSTM layers of 128 hidden states and a dropout rate of 4/10  followed by a dense layer of 128 connections. We attempted different state sizes but found 128 to be the smallest number with the best consistent performance.

We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure (X). The multimodal architecture allows our data to vary in size and given that all four levels of decomposition have varying length due to downsampling, we propose each model to learn the representation of each distinct feature. For this architecture we propose two models, a 1D CNN and an LSTM, in which we employ two CNN’s and 1 LSTM. They are all concatenated after their respective dense layer which performs the final   We tried numerous variations and saw no impact on performance.

\section{black box}
Our model is constructed of four different architectures, 2 LSTM and 2 CNN’s, that are concatenated at the dense layer. The four features we use are amplitude and phase of each level of decomposition for HV and VV. We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other. Wavelet decomposition produces a complex value but that is unusable in a standard neural network due to loss of information unless we move to a complex LSTM but as stated in {} there isn’t a significant increase in accuracy.

\section{white box}
Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method. We then pass those images through a CNN and those that are correctly classified are shown to the user.

In order to classify those images we created a CNN for it specifically and avoided using our existing CNN in order to make it deeper and increase performance when images aren’t that classifiable. We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it. We were able to achieve 60% accuracy on our dataset.

\section{Notation}

\section{Layout}

\chapter{Problem Statement}

\chapter{Related Work}

\chapter{Markov Transition Fields}
% We propose a framework similar to [] for encoding dynamical transition statistics,
% but we continue their work by considering ith order Markov transition probabilities.
% Given a time series X, we decompose its magnitude axis into two separate
% properties by representing it as a polar coordinate, Xangular and Xradial. We then
% identify the Q quantile bins for both properties and the temporal values Xtime
% and assign each x ∈ Xangular, Xradial, Xtime to its corresponding bin qj (j in
% [1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
% Wradial, Wtime, by counting transitions among quantile bins in the manner of a
% ith order Markov chain along the time axis.

% W does not take into account the temporal axis so to prevent any information
% loss we construct a Markov Transition Field (MTF) for each W. The MTF
% denotes the probability of transitioning from qi to qj for each x ∈ X. This,
% in turn, allows us to consider the transition probability on the magnitude and
% temporal axis.

% As described in [] the MTF encodes the multi-span transition probabilities
% of the time series, but given that we have three different M’s we modify their
% approach and consider each M to be a separate color channel of RGB where
% Mangular is red, Mtime is blue, Mradial is green. We perform linear interpolation
% on each M individually to extend the probabilities to be within the range of 0 and 255.

\chapter{Main Results}

\chapter{Conclusion}
\label{chap:conclusion}

% \bibliography{../references/main}
% \bibliographystyle{plain}

\end{document}

### 2020-03-12 paper/main.tex
new content

@@
 \chapter{Introduction}
-In this thesis, we study interpretable machine learning as applied to complex-valued radio signals.
+
+[explain what black box and white box is]
+In this thesis, we study interpretable machine learning as applied to complex-valued time series.
 Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.
-These methods, however, fall short of allowing users to visualize patterns within their dataset and .
-To address this issue, we propose a hybrid interpretable model that can be extended to any time series dataset.
-Our model consists of two models.
-The first is an ensemble method of classifiers that functions as our black box method.
-The second is a white-box model that encodes time series as images for each signal and outputs all images that have been correctly classified.
 
-We study the adaption of deep learning models to complex-valued radio signals.
-We compare the deep learning models such as CNN’s have become ever-increasingly powerful and widely used but their complex nature makes it difficult for users to understand what is going on so as black box models get more popular the need for an interpretable accomplice is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
+[add more reasons]
+[add how most time series are classified]
+These methods, however, fall short of allowing users to visualize patterns within their dataset and ().
 
-The growth of mobile devices and demand for wireless data has created a need for  High quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.
+To address this issue, we propose an interpretable hybrid  model that can be extended to any time series dataset.
+Our architecture is composed of two models, an ensemble method of classifiers that functions as a black box method, and a white-box model that encodes time series as images, attemps to classify them through a CNN, and outputs all images correctly classified in both black-box and white-box method.
+
+In our black-box model, we analyze  deep learning models such as CNN’s which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on.
+So, as black-box models get more popular the need for an interpretable accessory?
+is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
+
+\section{The need for interpretable models in management problems}
+The growth of mobile devices and demand for wireless data has created a need for igh quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.
 however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments.
-[7] [6] [9]
-This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically
+This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
 
-Building upon successful strategies from image and voice recognition domains in machine learning, we demonstrate an approach in the radio using Convolutional Neural Networks (CNNs) and Deep Neural Networks (DNNs) which offers flexibility to learn features across a wide range of tasks and demonstrates improved classification accuracy against current day approaches.
-Existing methods, however, fall short of allowing users to visualize patterns within their dataset.
-To address this issue, we propose a hybrid interpretable model that can be extended to any time series dataset.
-Our model consists of two models.
-The first is an ensemble method of classifiers that functions as our black box method.
-The second is a white-box model that encodes time series as images for each signal and outputs all images that have been correctly classified.
 \section{Contributions}
+[fill in]
+
+\section{Data}
+Our dataset contains two different types of signals, raw and discrete.
+The raw signal consists of time and amplitude with 31,000 data points while discrete data can vary from 22 to 44.
+In industry, they do not have access to raw data because it is difficult to track that information so we focus on discrete data for this thesis.
+
+The raw signal has the same time intervals across the entire dataset but it may have different start and end times.
+
+[ What is raw signal and discrete, compare them, ]
+
+\chapter{Preprocessing}
+We compared this method to other features extracted from Fourier-transform, Short Time Fourier transform, sliding window method, and polar coordinates taken directly from the signals amplitude/phase value.
+
+[move to results]
+Using any of those methods reduced our model's accuracy as well as increased training time.
+For Fourier Transform, we attempted to perform FT over the entire signal and use amplitude and time as a feature accuracy dropped below 90\%.
+
+We believe this is because FT only takes into account the frequency domain and ignores time.
+FFT and sliding window method gave us better performance averaging around 90\% accuracy.
+We also attempted to take the average of each window and tried different window sizes as well as window overlapping but did not see a significant improvement in loss or accuracy.
+[move to results]
+
+We normalize all data using L2 norm and format our data by each subsequent wavelet decomposition. For instance, if we do N levels of decomposition the Nth item would be at the end of the array.
+
+\section{What is it?}
+[what is feature extraction]
+
+Radio signal data is presented in a complex-value format which is unusable in a standard neural network so in an attempt to make our model robust and reduce any complexity we focused on using real values as features.
+
+% maybe?
+% Data pre-processing is concerned with the analysis and manipulation of the collected spectrum data to arrive at potentially good wireless data representations.
+The raw samples organized into data vectors rk in the previous block are pipelined as input for signal processing (SP) tools that analyze, process and transform the data to arrive at simple data representations such as frequency,
+% amplitude, phase and spectrum, or more complex features
+% xk such as e.g. cyclostationary features. In addition, feature
+% learning such as deep learning may be utilized to automatically
+% extract more low-level and high-level features. In many ML
+% applications the choice of features is just as important, if not
+% more important than the choice of the ML algorithm.
+
+\section{Transformations}
+
+\subsection{Transformation 1 STFT/ FT}
+% [Recall we can write a complex number in terms of its magnitude and phase
+% (i.e., its polar representation). ]
+
+Transformation 1 : STFT/ FT
+[what is it]
+
+[fft formula]
+The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions.
+Each basis function is a complex value of a frequency so it allows us to view our  data in frequency domain as opposed to amplitude.
+
+One of the shortcomings of the FT is rooted in the Heisenberg Uncertainty Principle (HUP) [quote?].
+The HUP states that the position and velocity of an object cannot be simulteanously measured which can be applied to the time frequency information of a signal.
+This means we cannot know which spectral components exist at any given time.
+The closest we can get is sampling at different ranges of time and finding a range of frequencies within that time frame.
+This method is described as the Short Time Fourier Transform.
+
+% The Fourier Transform is a magical mathematical tool.
+The Fourier Transform decomposes any function into a sum of sinusoidal basis functions.
+Each of these basis functions is a complex exponential of a different frequency.
+The Fourier Transform, therefore gives us a unique way of viewing any function - as the sum of simple sinusoids.
+
+% The Fourier Series showed us how to rewrite any periodic function into a sum of sinusoids. The Fourier Transform is the extension of this idea to non-periodic functions.
+
+% While the Fourier Transform is a beautiful mathematical tool, its widespread popularity is due to its practical application in virtually every field of science and engineering.
+It's hard to understand why the Fourier Transform is so important.
+But I can assure you it enables the solution to difficult problems be made simpler (and also makes previously unsolved problems solvable).
+In addition, the Fourier Transform gives us a new method of viewing the world, which is fantastic for giving a more intuitive feel for our universe.
+
+\subsection{Shortcoming stft}
+[stft]
+The Short-Time Fourier transform (STFT), is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.[Ref] 
+Computing STF requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
+This allows us to view the the fourier spectrum at a more granular level which could potentially reveal different patterns amongst signals of different classes.
+
+[why it performed poorly]
+One of the issues with this method is we can create very narrow window sizes which gives us a better understanding of the data with respect to time but we lose understanding of the frequency domain.
+Additionally, selecting an appropriate window size for segementing the signal can be an arduous that would require fine-tuning as well as well as increasing the dimensionality of our dataset.
+
+% The problem with STFT is the fact whose roots go back to what is known as the Heisenberg
+% Uncertainty Principle. This principle originally applied to the momentum and location of
+% moving particles, can be applied to time-frequency information of a signal. Simply, this principle
+% states that one cannot know the exact time-frequency representation of a signal, i.e., one cannot
+% know what spectral components exist at what instances of times. What one can know are the
+% time intervals in which certain bands of frequencies exist, which is a resolution problem.
+% The problem with the STFT has something to do with the width of the window function that is
+% used. To be technically correct, this width of the window function is known as the support of
+% the window. If the window function is narrow, then it is known as compactly supported. This
+% terminology is more often used in the wavelet world, as we will see later.
+
+% The problem is that when you are studying a real signal it would be useful to know what the ‘instantaneous frequency’ of the signal is.
+The instantaneous frequency is the exact frequency of a signal at an exact moment in time.
+For instance if I was listening to a music track I would like to be able to say ‘at 1 min 59.0423 seconds into the music track the so,und is 1563.2 Hz’.
+Unfortunately the Fourier Transform cannot do this because there exists a minimum amount of uncertainty between the frequency and time domains, like Heisenberg has a minimum amount of uncertainty between the speed and position of a particle (the area of the box).
+You can know the moment in time you want to find the frequency for (like the blue box), but because there is a minimum uncertainty, the box of has to stretch out across frequency, meaning that you are unsure of the frequency at that moment in time.
+
+% The best you can do with a Fourier Transform is to sample a range of time (for instance, the time signal between 1 min 58 sec and 1 min 59 sec in a song) and find a range of frequencies that were played over that amount of time, as represented by the black box.
+An example of this can be seen by looking at the third picture in the post again.
+There is a signal over a range of time (e.g.
+someone saying ‘hello’) and the frequency graph is the range of frequencies recorded over that time.
+
+\subsection{Transformation 2: Abs/angle}
+
+Transformation 2: Amplitude / Phase
+The representation of a complex number as a sum of a real and imaginary number, z = x + iy, is called its Cartesian representation.
+For every cartesian point, we calculate its radius and angular distance which are real values and use those as features.
+
+[purpose and motivation? Behind making it ang/abs]
+
+[definition plus method]
+
+\subsection{Transformation 2: Sliding window }
+
+Transformation 2: Sliding window 
+We propose a method similar to STFT but in an effort to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
+
+\subsection{Transformation 3: Wavelets}
+Transformation 3: Wavelets
+
+Wavelets are similar to the FT in which they deconstruct a signal using representations of other signals.
+The key difference is wavelet signals are finite in time and frequency as opposed to sin and cos signals which can carry on forever.
+Simply, this allows us to extract information from a signal which respect to time and location.
+
+The use of wavelets is called the wavelet transform which is a technique in wich a signal is analyzed using different versions of a dilated and translated basis functionss called the mother wavelet.
+There are two types of wavelet transformations, discrete and continous.
+In this thesis we focus on discrete.
+
+DWT uses a discrete set of wavelet scales and translations which decompose the signal into mutually orthogional set of wavelets.
+
+One advantage of wavelets is they have varying window sizes so they can be wide for slow frequencies and narrow for faster ones this is handled by the property called scale [add figure].
+
+wavelets
+
+Some properties of wavelets are orthoganility, zero mean, and are efficient at representing localized data and functions.
+
+Db4 wavelet http://wavelets.pybytes.com/wavelet/db4/
+
+Model 1:
+
+Model 2: 
+Model 3: `'
+
+In the preprocessing stage we tried numerous combinations of features but found the best performing features to come from wavelet decomposition.
+
+Wavelets
+
+We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank which means it’s composed of cascading high-pass and low-pass filters.
+This gives us the advantage of splitting a signal into several frequency sub-bands.
+Wavelet decompositions give us the advantage of gaining features that take into account frequency and time domains.
+
+[explain why they performed better]
+
+[alternatives]
+
+\section{Data Shape}
+[Shape of data comparison]
+
+A neural network requires all data to be the same shape but because we downsample in wavelet decomposition, each data size is halved until making it impossible to place all levels of decomposition into the same dataset.
+In order to circumvent this we tried two different methods, resampling the data, and treating each level of decomposition as its own dataset.
+
+\subsection{Resampling}
+
+We use spline interpolation in order to resize each level of decomposition into a single array.
 
 \section{Models}
-Our black box method is composed of two types of neural networks, Convolutional Neural Networks and Long Short Term Memory networks.
+[introduce models definitions]
+[EXPLAIN LOSS AND HOW ACCURACY IS MEASURED]
 
-CNN
+This section gives an overview of the theoretical background of concepts discussed in this work.
+
+Machine learning algorithms can simply be seen as learning a target function (f) that maps input data (X) to an output (Y).
+There are several techniques to make this work but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short Term Memory networks (LSTM).
+Nonparametric algorithms are algorithms that attempt to make minimal assumptions about the form of the function so they can learn any form from data provided to it.
+An example would be a neural network because it has no prior knowledge about what it’s classifying and attempts to generalize any new data points that it has not seen before.
+
+[cnn]
+
+A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis.
+It consists of an input and output layer as well as hidden layers.
+A convolution is an operation between a vector of weights w against an input x.
+It consists of taking the dot product between m and x were in steps of a filter size defined by n.
+
+Convolutional neural networks perform feature learning via non-linear transformations implemented as a series of nested layers.
+The input data is a multidimensional data array, called tensor, that is presented at the visible layer.
+This is typically a grid-like topological structure, e.g.
+time-series data, which can be seen as a 1D grid taking samples at regular time intervals, pixels in images with a 2D layout, a 3D structure of videos, etc.
+Then a series of hidden layers extract several abstract features.
+Those layers are “hidden” because their values are not given.
+Instead, the deep learning model must determine which data representations are useful for explaining the relationships in the observed data.
+Each layer consists of several kernels that perform a convolution over the input; therefore, they are also referred to as convolutional layers.
+Kernels are feature detectors, that convolve over the input and produce a transformed version of the data at the output.
+Those are banks of finite impulse response filters as seen in signal processing, just learned on a hierarchy of layers.
+The filters are usually multidimensional arrays of parameters that are learnt by the learning algorithm [24] through a training process called backpropagation.
+
+[lstm]
+Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture[1] used in the field of deep learning.
+Unlike standard feedforward neural networks, LSTM has feedback connections.
+It can not only process single data points (such as images), but also entire sequences of data (such as speech or video).
+For example, LSTM is applicable to tasks such as unsegmented, connected handwriting recognition[2], speech recognition[3][4] and anomaly detection in network traffic or IDS's (intrusion detection systems).
+
+A common LSTM unit is composed of a cell, an input gate, an output gate and a forget gate.
+The cell remembers values over arbitrary time intervals and the three gates regulate the flow of information into and out of the cell.
+
+LSTM networks are well-suited to classifying, processing and making predictions based on time series data, since there can be lags of unknown duration between important events in a time series.
+LSTMs were developed to deal with the exploding and vanishing gradient problems that can be encountered when training traditional RNNs.
+Relative insensitivity to gap length is an advantage of LSTM over RNNs, hidden Markov models and other sequence learning methods in numerous applications.
+
+[multimodal]
+Due to the superior performance and computationally tractable representation capability (in vector spaces) in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks [38, 39, 44].
+Typically, domain-specific neural networks are used on different modalities to generate their representations and the individual representations are merged or aggregated.
+Finally, the prediction is made on top of aggregated representation usually with another a neural network to capture the interactions between modalities and learn complex function mapping between input and output.
+Addition (or average) and concatenation are two common aggregation methods, i.e.,
+The network structure is illustrated in Figure 1(b) .
+The arrows are function mappings or computing operations.
+The dotted boxes are representations of single and combined modality features.
+We call them additive combinations because their critical step is to add modality hidden vectors (although often in a nonlinear way).
+
+[explain architectures]
+
+\subsection{CNN}
+
 Our CNN has an architecture consisting of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X).
@@
 
-LSTM
+\subsection{LSTM}
 We wanted to keep our LSTM network as small as possible, for X purposes, so we went with two LSTM layers of 128 hidden states and a dropout rate of 4/10  followed by a dense layer of 128 connections.
 We attempted different state sizes but found 128 to be the smallest number with the best consistent performance.
+
+We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure (X).
+The multimodal architecture allows our data to vary in size and given that all four levels of decomposition have varying length due to downsampling, we propose each model to learn the representation of each distinct feature.
+For this architecture we propose two models, a 1D CNN and an LSTM, in which we employ two CNN’s and 1 LSTM.
+They are all concatenated after their respective dense layer which performs the final   We tried numerous variations and saw no impact on performance.
 
@@
 We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it.
-We were able to achieve 60% accuracy on our dataset.
+We were able to achieve 60\% accuracy on our dataset.
 

### 2020-03-12 paper/main.tex
deleting code files

@@
 [ What is raw signal and discrete, compare them, ]
-
+[explain hv, vv]
 \chapter{Preprocessing}
@@
 We use spline interpolation in order to resize each level of decomposition into a single array.
+We tested different sizes of interpolation from N to N/3 where N is the size of the largest discrete signal size and found no drop in performance.
 
-\section{Models}
+\subsection{Each level}
+Figure {X} depicts our model in which each level of decomposition is treated as its own dataset. We found no difference in accuracy compared to the resampling method.
+
+\chapter{Models}
 [introduce models definitions]
@@
 
+\section{Definition}
 This section gives an overview of the theoretical background of concepts discussed in this work.
@@
 
+\subsection{CNN}
 [cnn]
@@
 
+\subsection{LSTM}
+
 [lstm]
@@
 
+\subsection{multimodal}
+
 [multimodal]
@@
 
+\section{Architectures Used}
+
 \subsection{CNN}
 
-Our CNN has an architecture consisting of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X).
-We found no improvements in performance when we attempted to make it deeper by adding additional convolutional layers as well as skip connections.
-
+Our CNN architecture consists of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X).
+We attempted to make it deeper but found inconsistent performance across all of our datasets.
+[add figure]
 \subsection{LSTM}
@@
 
+[add figure]
+
+\subsection{Final method}
 We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure (X).
@@
 They are all concatenated after their respective dense layer which performs the final   We tried numerous variations and saw no impact on performance.
+[add figure]
 
@@
 We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it.
-We were able to achieve 60\% accuracy on our dataset.
+
+\subsection{mtf}
+We approach our MTF is slightly different than [] because instead of performing it over all samples in the dataset we perform it for every signal. The three attributes we use are values o
 
@@
 \chapter{Markov Transition Fields}
-% We propose a framework similar to [] for encoding dynamical transition statistics,
-% but we continue their work by considering ith order Markov transition probabilities.
-% Given a time series X, we decompose its magnitude axis into two separate
-% properties by representing it as a polar coordinate, Xangular and Xradial. We then
-% identify the Q quantile bins for both properties and the temporal values Xtime
-% and assign each x ∈ Xangular, Xradial, Xtime to its corresponding bin qj (j in
-% [1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
-% Wradial, Wtime, by counting transitions among quantile bins in the manner of a
-% ith order Markov chain along the time axis.
+We propose a framework similar to [] for encoding dynamical transition statistics,
+but we continue their work by considering ith order Markov transition probabilities.
+Given a time series X, we decompose its magnitude axis into two separate
+properties by representing it as a polar coordinate, Xangular and Xradial. We then
+identify the Q quantile bins for both properties and the temporal values Xtime
+and assign each x in Xangular, Xradial, Xtime to its corresponding bin qj (j in
+[1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
+Wradial, Wtime, by counting transitions among quantile bins in the manner of a
+ith order Markov chain along the time axis.
 
-% W does not take into account the temporal axis so to prevent any information
-% loss we construct a Markov Transition Field (MTF) for each W. The MTF
-% denotes the probability of transitioning from qi to qj for each x ∈ X. This,
-% in turn, allows us to consider the transition probability on the magnitude and
-% temporal axis.
+W does not take into account the temporal axis so to prevent any information
+loss we construct a Markov Transition Field (MTF) for each W. The MTF
+denotes the probability of transitioning from qi to qj for each x in X. This,
+in turn, allows us to consider the transition probability on the magnitude and
+temporal axis.
 
-% As described in [] the MTF encodes the multi-span transition probabilities
-% of the time series, but given that we have three different M’s we modify their
-% approach and consider each M to be a separate color channel of RGB where
-% Mangular is red, Mtime is blue, Mradial is green. We perform linear interpolation
-% on each M individually to extend the probabilities to be within the range of 0 and 255.
+As described in [] the MTF encodes the multi-span transition probabilities
+of the time series, but given that we have three different M’s we modify their
+approach and consider each M to be a separate color channel of RGB where
+Mangular is red, Mtime is blue, Mradial is green. We perform linear interpolation
+on each M individually to extend the probabilities to be within the range of 0 and 255.
+
+\section{}
 

### 2020-03-12 code/collinear.py

@@ -1,15 +1,0 @@
-import numpy as np
-from shapes import random_shape
-
-def random_example():
-    points = np.random.random_sample((3, 2))
-    speeds = np.random.random_sample((3, 2))
-
-    speeds = np.linalg.norm(speeds, axis=0) / 1e-3
-    
-    positions = np.arange(1, 1001)[:,np.newaxis] * speeds 
-
-    print(positions)
-
-if __name__ == '__main__':
-    random_example()

### 2020-03-12 code/formation.py

@@ -1,202 +1,0 @@
-import numpy as np 
-import math
-
-import matplotlib.pyplot as plt
-
-from functools import partial 
-from itertools import permutations, combinations
-from typing import Callable, Dict, Tuple, List
-
-from plotters import *
-
-def get_transform(v1: np.array, v2: np.array) -> float:
-    norm_1 = np.linalg.norm(v1)
-    norm_2 = np.linalg.norm(v2)
-    angle = math.acos(np.dot(v1, v2) / (norm_1 * norm_2))
-    scale = norm_2 / norm_1
-
-    if np.linalg.det([v1.astype(np.float32), v2.astype(np.float32)]) > 0:
-        angle = -angle
-
-    return angle, scale
-
-def replication(formation: np.array, 
-                q_0: np.array, q_1: np.array, 
-                i: int=0, j: int=1) -> np.array:
-    """Returns the replicated shape.
-
-    This returns the shape P similar to formation such that
-    P's ith and jth points are fixed to q_0 and q_1, respectively.
-
-    Args:
-        formation (np.array): formation to replicate
-        q_0: Point to fix ith point in target formation to
-        q_1: Point to fix jth point in target formation to
-        i: index of point in target formation to fix to q_0
-        j: index of point in target formation to fix to q_1
-
-    Returns:
-        np.array: The replicated shape
-    """
-    formation = np.copy(formation)
-    q_0 = np.copy(q_0)
-    q_1 = np.copy(q_1)
-
-    assert(len(formation.shape) == 2)   # two-dimensional
-    assert(formation.shape[1] == 2)     # collection of two-dimensional points
-    assert(q_0.shape == (2,))           # two-dimensional point
-    assert(q_1.shape == (2,))           # two-dimensional point
-
-    formation -= (formation[i] - q_0)   # shift to i to q_0
-
-    offset = np.copy(formation[i])
-    formation -= offset     # temp shift i to origin
-    q_1 -= offset
-
-    theta, scale = get_transform(formation[j], q_1)
-
-    c = math.cos(-theta)
-    s = math.sin(-theta)
-
-    formation = np.array(
-        [
-            formation[:,0] * c - formation[:,1] * s,
-            formation[:,0] * s + formation[:,1] * c
-        ]
-    ).T * scale + offset
-
-    return formation
-
-def replication_machine_circles(formation: np.array,  
[…]

### 2020-03-12 code/gabriel.py

@@ -1,75 +1,0 @@
-import numpy as np
-import math
-from itertools import product, chain, combinations, permutations, starmap
-from functools import partial
-
-def dist(p1, p2):
-    return math.sqrt((p1[0] - p2[0])**2 + (p1[1] - p2[1])**2)
-
-def mid(p1, p2):
-    return (p1[0] + p2[0]) / 2, (p1[1] + p2[1]) / 2
-
-def get_gabriel_graph(points: np.array, points2: np.array = None):
-    if points2 is None:
-        points2 = points
-    aux = []
-    for p1, p2 in product(points, points2):
-        if p1 == p2:
-            continue
-        m = mid(p1, p2)
-        aux.append((m, dist(m, p1), p1, p2))
-
-    edges = []
-    for m, rad, p1, p2 in aux:
-        if (True in map(lambda p: dist(m, p) < rad, points) and
-            True in map(lambda p: dist(m, p) < rad, points2)):
-            continue
-        edges.append((p1, p2))
-
-    return edges
-
-def brute_match(points, points2):
-    edges = []
-    r = None
-    for _points2 in permutations(points2):
-        _r = max(starmap(dist, zip(points, _points2)))
-        if r is None or _r < r:
-            r = _r
-            edges = list(zip(points, _points2))
-    return edges
-
-import networkx as nx
-import matplotlib.pyplot as plt
-from matplotlib import collections  as mc
-
-if __name__ == '__main__':
-    points = list(map(tuple, np.random.rand(5, 2)))
-    points2 = list(map(tuple, np.random.rand(5, 2)))
-    edges = set(get_gabriel_graph(points, points2))
-
-    print('Checking Brutely')
-    brute_edges = set(brute_match(points, points2))
-    if brute_edges.issubset(edges):
-        print('TRUTH')
-    else:
-        print('FALSE')
-
-    fig: plt.Figure
-    ax: plt.Axes
-    fig, ax = plt.subplots()
-    ax.scatter(*zip(*points), color='b')
-    ax.scatter(*zip(*points2), color='y')
-
-    lc = mc.LineCollection(edges)
-    ax.add_collection(lc)
-    
-    lc_brute = mc.LineCollection(brute_edges, color='g')
-    ax.add_collection(lc_brute)
-    
-    lc_absent = mc.LineCollection(brute_edges - edges, color='r')
-    ax.add_collection(lc_absent)
-
-    ax.autoscale()
-    ax.margins(0.1)
-    plt.show()
-

### 2020-03-12 code/geometry.py

@@ -1,39 +1,0 @@
-import numpy as np
-from math import sqrt
-
-def circle_intersection(circle1, circle2):
-        '''
-        @summary: calculates intersection points of two circles
-        @param circle1: tuple(x,y,radius)
-        @param circle2: tuple(x,y,radius)
-        @result: tuple of intersection points (which are (x,y) tuple)
-        '''
-        # return self.circle_intersection_sympy(circle1,circle2)
-        x1,y1,r1 = circle1
-        x2,y2,r2 = circle2
-        # http://stackoverflow.com/a/3349134/798588
-        dx,dy = x2-x1,y2-y1
-        d = sqrt(dx*dx+dy*dy)
-        if d > r1+r2:
-            print("#1")
-            return None # no solutions, the circles are separate
-        if d < abs(r1-r2):
-            print("#2")
-            return None # no solutions because one circle is contained within the other
-        if d == 0 and r1 == r2:
-            print("#3")
-            return None # circles are coincident and there are an infinite number of solutions
-
-        a = (r1*r1-r2*r2+d*d)/(2*d)
-        h = sqrt(r1*r1-a*a)
-        xm = x1 + a*dx/d
-        ym = y1 + a*dy/d
-        xs1 = xm + h*dy/d
-        xs2 = xm - h*dy/d
-        ys1 = ym - h*dx/d
-        ys2 = ym + h*dx/d
-
-        return np.array([
-            [xs1, ys1],
-            [xs2, ys2]
-        ])

### 2020-03-12 code/helpers.py

@@ -1,19 +1,0 @@
-import numpy as np
-
-def get_intersect(a1, a2, b1, b2):
-    """ 
-    Returns the point of intersection of the lines passing through a2,a1 and b2,b1.
-    a1: [x, y] a point on the first line
-    a2: [x, y] another point on the first line
-    b1: [x, y] a point on the second line
-    b2: [x, y] another point on the second line
-    """
-    s = np.vstack([a1, a2, b1, b2])        # s for stacked
-    h = np.hstack((s, np.ones((4, 1))))    # h for homogeneous
-    l1 = np.cross(h[0], h[1])              # get first line
-    l2 = np.cross(h[2], h[3])              # get second line
-    x, y, z = np.cross(l1, l2)             # point of intersection
-    if z == 0:                             # lines are parallel
-        return np.array([float('inf'), float('inf')])
-    return np.array([x/z, y/z])
-

### 2020-03-12 code/optimization.py

@@ -1,169 +1,0 @@
-from cvxopt import matrix, solvers
-from math import sqrt 
-import numpy as np
-import matplotlib.pyplot as plt
-from itertools import permutations, combinations, chain
-from plotters import get_colors
-
-from typing import List
-
-from formation import triangle_solution, triangle_solution_no_perm, replication
-from shapes import random_shape, regular_shape, Triangle
-solvers.options['show_progress'] = False
-
-def print_arrays(**kwargs):
-    name: str
-    value: np.array
-    for name, value in kwargs.items():
-        print(f'{name} {value.shape}')
-        print(value)
-        print()
-
-def format_eqs(lefts: List[List[float]], right: List[float]):
-    *xs, cs = zip(*([right] + lefts))    
-    mat = []
-    for elems in xs:
-        mat.append([float(-elem) for elem in elems])
-
-    return matrix(mat), matrix(cs)
-
-import math
-def convex_solution_no_perm(positions: np.array, formation: np.array, verbose: bool = False) -> np.array:
-    positions = np.copy(positions)
-    formation = np.copy(formation) - formation[0] # adjust formation to origin
-    angle = -math.atan2(formation[1][1], formation[1][0])
-    xs, ys = formation[:,0].copy(), formation[:,1].copy()
-    formation[:,0] = math.cos(angle) * xs - math.sin(angle) * ys
-    formation[:,1] = math.sin(angle) * xs + math.cos(angle) * ys
-
-    # assert((formation[0] == np.array([0.0, 0.0])).all())
-    assert(positions.shape == formation.shape)
-
-    s2norm = np.linalg.norm(formation[1])
-    n = len(formation)
-
-    # | q_i.x - p_i.x |    
-    # |               |     <= r
-    # | q_i.y - p_i.y |_2 
-    # 
-    # such that 
-    # Aq = 0    <---    this says the shape is similar
-
-    Gs = []
-    hs = []
-
-    eqs = np.append(np.eye(2*n), np.zeros((2*n, 2)), axis=1)
-    eqs[:,-1][::2] = -positions[:,0]
-    eqs[:,-1][1::2] = -positions[:,1]
-
-    eq_r = np.zeros(2*n + 2)
-    eq_r[-2] = 1.0
-    for i in range(n):
-        G, h = format_eqs(
-            [eqs[2*i].tolist(), eqs[2*i+1].tolist()],
[…]

### 2020-03-12 code/plotters.py

@@ -1,33 +1,0 @@
-
-import matplotlib.pyplot as plt
-import matplotlib.cm as cm
-import numpy as np 
-
-from typing import Iterable
-
-def get_colors(num: int) -> Iterable[str]:
-    return cm.rainbow(np.linspace(0, 1, num))
-
-def plot_shape(ax: plt.Axes, shape: np.array, color: str = 'black', marker: str = 'o') -> None:
-    ax.scatter(x=shape[:,0], y=shape[:,1], color=color, marker=marker)
-
-def plot_shapes(ax: plt.Axes, *shapes: np.array, marker: str = 'o') -> None:
-    colors = get_colors(len(shapes))
-    for shape, c in zip(shapes, colors):
-        plot_shape(ax, shape, color=c, marker=marker)
-
-def plot_polygon(ax: plt.Axes, polygon: np.array, color: str = 'black') -> None:
-    ax.add_patch(plt.Polygon(polygon, color=color, fill=False))
-
-def plot_polygons(ax: plt.Axes, *polygons: np.array) -> None:
-    colors = get_colors(len(polygons))
-    for polygon, color in zip(polygons, colors):
-        plot_polygon(ax, polygon, color=color)
-
-def plot_circle(ax: plt.Axes, point: np.array, radius: float, color: str = 'black') -> None:
-    ax.add_patch(plt.Circle(point, radius, color=color, fill=False))
-
-def plot_circles(ax: plt.Axes, points: np.array, radiuses: np.array) -> None:
-    colors = get_colors(len(points))
-    for point, radius, color in zip(points, radiuses, colors):
-        plot_circle(ax, point, radius, color=color)

### 2020-03-12 code/shapes.py

@@ -1,39 +1,0 @@
-import numpy as np 
-import math 
-
-def random_shape(num: int, low: float = -1.0, high: float = 1.0) -> np.ndarray:
-    return np.random.random_sample((num, 2)).astype(np.longdouble) * (high - low) + low
-
-def regular_shape(num: int) -> np.ndarray:
-    return np.array(
-        [
-            [math.cos(2 * math.pi * i / num), math.sin(2 * math.pi * i / num)]
-            for i in range(num)
-        ]
-    ).astype(np.longdouble)
-
-class Triangle:
-    EQUILATERAL = np.array([
-        [0.0, 0.0],
-        [1.0, 0.0],
-        [0.5, math.sqrt(3) / 2]
-    ]).astype(np.longdouble)
-    
-    @staticmethod
-    def right(angle: float, deg: float = False) -> np.ndarray:
-        if deg:
-            angle = angle * math.pi / 180
-        return np.array([
-            [0.0, 0.0],
-            [1.0, 0.0], 
-            [1.0, math.tan(angle)]
-        ]).astype(np.longdouble)
-
-    @staticmethod
-    def isosoles(height: float) -> np.ndarray:
-        return np.array([
-            [-0.5, 0.0],
-            [0.0, height],
-            [0.5, 0.0],
-        ]).astype(np.longdouble)
-

### 2020-03-12 code/solution.py

@@ -1,129 +1,0 @@
-from formation import triangle_solution_no_perm, replication
-from itertools import product, permutations, combinations
-
-import numpy as np
-
-def is_similar(shp, _shp):
-    sides = np.linalg.norm(shp - np.roll(shp, 1, axis=0), axis=1)
-    sides =  np.sort(sides) / np.sum(sides)
-    _sides = np.linalg.norm(_shp - np.roll(_shp, 1, axis=0), axis=1)
-    _sides = np.sort(_sides) / np.sum(_sides)
-    return np.isclose(sides, _sides).all()
-
-def solution(positions: np.array, formation: np.array) -> np.array:
-    combos = lambda x: map(np.array, combinations(range(x), 3))
-    n = len(positions)
-    target, radius = None, None
-    for i, j in product(combos(n), combos(n)):
-        pos = positions[i]
-        for perm in map(np.array, permutations(range(3))):
-            tri_target, tri_radius = triangle_solution_no_perm(pos, formation[j[perm]])
-                
-            _target = replication(formation, tri_target[0], tri_target[1], j[perm][0], j[perm][1])
-
-            tmp_radius = None
-            for _tar in map(np.array, permutations(_target)):
-                _radiuses = np.linalg.norm(positions - _tar, axis=1)
-                _radius = np.max(_radiuses)
-                if radius is None or _radius < radius:
-                    target, radius = _tar, _radius
-                if tmp_radius is None or _radius < tmp_radius:
-                    tmp_radius = _radius
-    return target
-
-from shapes import random_shape
-from optimization import convex_solution, convex_solution_no_perm
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-
-# if __name__ == '__main__':
-#     n, m = 50, 20
-#     data = np.zeros((n, m))
-#     for i in range(3, n):
-#         total = 0
-#         for j in range(m):
-#             positions, formation = random_shape(i), random_shape(i)
-#             target, radius = convex_solution(positions, formation)
-#             radiuses = np.linalg.norm(target - positions, axis=1)
-#             num = len(radiuses[np.isclose(radiuses, radius)])
-#             print(f'{i}\t{j}\t- {num}')
[…]

### 2020-03-12 code/tests.py

@@ -1,436 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import math
-from itertools import combinations, product
-from matplotlib.animation import FuncAnimation
-from matplotlib.widgets import Slider, Button, RadioButtons
-from matplotlib.collections import PatchCollection
-
-from multiprocessing import Pool
-from functools import partial
-import signal
-
-from formation import (
-    replication,
-    replication_machine,
-    replication_machine_circles,
-    replication_spanner,
-    replication_spanner_circles,
-    triangle_solution,
-    triangle_solution_no_perm,
-    approximation,
-    get_sides
-)
-from geometry import circle_intersection
-from plotters import *
-from shapes import Triangle, random_shape, regular_shape
-from typing import Tuple, List, Optional
-
-# https://stackoverflow.com/questions/3252194/numpy-and-line-intersections/3252222
-def get_intersect(a1, a2, b1, b2):
-    """
-    Returns the point of intersection of the lines passing through a2,a1 and b2,b1.
-    a1: [x, y] a point on the first line
-    a2: [x, y] another point on the first line
-    b1: [x, y] a point on the second line
-    b2: [x, y] another point on the second line
-    """
-    s = np.vstack([a1, a2, b1, b2])        # s for stacked
-    h = np.hstack((s, np.ones((4, 1))))    # h for homogeneous
-    l1 = np.cross(h[0], h[1])              # get first line
-    l2 = np.cross(h[2], h[3])              # get second line
-    x, y, z = np.cross(l1, l2)             # point of intersection
-    if z == 0:                             # lines are parallel
-        return np.array([float('inf'), float('inf')])
-    return np.array([x / z, y / z])
-
-
-def test_same_solution(positions: Optional[np.array] = None,
-                       formation: Optional[np.array] = None,
-                       speed: float = 0.002):
-    if positions is None:
-        positions = random_shape(3)
-        print(positions)
-    if formation is None:
-        formation = regular_shape(3)
-
-    fig: plt.Figure
-    ax: plt.Axes
-    fig, ax = plt.subplots()
-    plt.subplots_adjust(bottom=0.25)
-    ax.axis('equal')
-    
-    axslider = plt.axes([0.25, 0.1, 0.65, 0.03], facecolor='lightgoldenrodyellow')
[…]

### 2020-03-13 paper/main.tex
more text

@@
 
-   \textbf{hybrid model}
+   \textbf{Hybrid Model for Interpretable Time Series Analysis}
 
@@
 \section{Data}
-Our dataset contains two different types of signals, raw and discrete.
-The raw signal consists of time and amplitude with 31,000 data points while discrete data can vary from 22 to 44.
-In industry, they do not have access to raw data because it is difficult to track that information so we focus on discrete data for this thesis.
 
-The raw signal has the same time intervals across the entire dataset but it may have different start and end times.
+We analyzed our model on three radio signal datasets generated using Spatial Modulation.
+Spatial Modulation is a transmission technique that uses multiple antennas.
+It maps a block of information bits to two units, one is a symbol chosen from a constellation diagram, and the second one is a unique transmit antenna number that is chosen from a set of transmit antennas.
+The method in which transmit antennas send and receive radio waves can be described through polarization.
+Two popular basis polarizations are horizontal linear, H, and vertical linear, V.
+Our datasets contain data horizontally transmitted and vertically received (HV) and data vertically transmitted and vertically received.
+
+The two properties HV and VV, are given to us in raw discrete format.
+Both properties are consist of time and amplitude values but vary in length and in how they are processed.
+The raw signal consists of over 31,000 data points with the exact time it was received.
+The size of the discrete data can vary from 22 to 44 data points, and is a processed representation of the raw signal that has equally spaced values.
 
@@
 \chapter{Preprocessing}
-We compared this method to other features extracted from Fourier-transform, Short Time Fourier transform, sliding window method, and polar coordinates taken directly from the signals amplitude/phase value.
+Radio signal data is presented in a complex-valued format that is unusable in a typical neural network, so in an attempt to make our model robust and reduce any complexity, we focused on using real values as features.
 
-[move to results]
-Using any of those methods reduced our model's accuracy as well as increased training time.
-For Fourier Transform, we attempted to perform FT over the entire signal and use amplitude and time as a feature accuracy dropped below 90\%.
+We tried numerous methods to extract real-valued features, including Fourier Transform, Short Time Fourier Transform, a custom sliding window method, and polar coordinates taken directly from the signals amplitude/phase value.
+Before applying any of these methods, we normalized all data using L2 normalization.
 
-We believe this is because FT only takes into account the frequency domain and ignores time.
-FFT and sliding window method gave us better performance averaging around 90\% accuracy.
-We also attempted to take the average of each window and tried different window sizes as well as window overlapping but did not see a significant improvement in loss or accuracy.
-[move to results]
+% \section{What is it?}
+% [what is feature extraction]
 
-We normalize all data using L2 norm and format our data by each subsequent wavelet decomposition. For instance, if we do N levels of decomposition the Nth item would be at the end of the array.
-
-\section{What is it?}
-[what is feature extraction]
-
-Radio signal data is presented in a complex-value format which is unusable in a standard neural network so in an attempt to make our model robust and reduce any complexity we focused on using real values as features.
-
-% maybe?
-% Data pre-processing is concerned with the analysis and manipulation of the collected spectrum data to arrive at potentially good wireless data representations.
+% % maybe?
+% % Data pre-processing is concerned with the analysis and manipulation of the collected spectrum data to arrive at potentially good wireless data representations.
 The raw samples organized into data vectors rk in the previous block are pipelined as input for signal processing (SP) tools that analyze, process and transform the data to arrive at simple data representations such as frequency,
-% amplitude, phase and spectrum, or more complex features
-% xk such as e.g. cyclostationary features. In addition, feature
-% learning such as deep learning may be utilized to automatically
-% extract more low-level and high-level features. In many ML
-% applications the choice of features is just as important, if not
-% more important than the choice of the ML algorithm.
+% % amplitude, phase and spectrum, or more complex features
+% % xk such as e.g. cyclostationary features. In addition, feature
+% % learning such as deep learning may be utilized to automatically
+% % extract more low-level and high-level features. In many ML
+% % applications the choice of features is just as important, if not
+% % more important than the choice of the ML algorithm.
 
@@
 
-\subsection{Transformation 1 STFT/ FT}
+\subsection{STFT/ FT}
 % [Recall we can write a complex number in terms of its magnitude and phase
@@
 
-Transformation 1 : STFT/ FT
+% Transformation 1 : STFT/ FT
 [what is it]
@@
 The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions.
-Each basis function is a complex value of a frequency so it allows us to view our  data in frequency domain as opposed to amplitude.
+Each basis function is a complex value of a frequency, so it allows us to view our data in the frequency domain as opposed to amplitude.
 
 One of the shortcomings of the FT is rooted in the Heisenberg Uncertainty Principle (HUP) [quote?].
-The HUP states that the position and velocity of an object cannot be simulteanously measured which can be applied to the time frequency information of a signal.
+The HUP states that the position and velocity of an object cannot be simultaneously measured, which can be applied to the time-frequency information of a signal.
 This means we cannot know which spectral components exist at any given time.
 The closest we can get is sampling at different ranges of time and finding a range of frequencies within that time frame.
-This method is described as the Short Time Fourier Transform.
+This method is described as the Short-Time Fourier Transform.
 
@@
 [stft]
-The Short-Time Fourier transform (STFT), is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.[Ref] 
+The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.[Ref] 
 Computing STF requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
-This allows us to view the the fourier spectrum at a more granular level which could potentially reveal different patterns amongst signals of different classes.
+This allows us to view the Fourier spectrum at a more granular level, which could potentially reveal different patterns amongst signals of different classes.
 
-[why it performed poorly]
-One of the issues with this method is we can create very narrow window sizes which gives us a better understanding of the data with respect to time but we lose understanding of the frequency domain.
-Additionally, selecting an appropriate window size for segementing the signal can be an arduous that would require fine-tuning as well as well as increasing the dimensionality of our dataset.
+One of the issues with this method is that we can create very narrow window sizes, which gives us a better understanding of the data with respect to time, but we lose understanding of the frequency domain.
+Additionally, selecting an appropriate window size for segmenting the signal can be an arduous task that would require fine-tuning as well as increasing the dimensionality of our dataset.
 
@@
 
-\subsection{Transformation 2: Abs/angle}
+\subsection{Polar Coordinates}
 
@@
 Transformation 2: Sliding window 
-We propose a method similar to STFT but in an effort to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
+We propose a method similar to STFT but to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
 
-\subsection{Transformation 3: Wavelets}
-Transformation 3: Wavelets
+\subsection{Wavelets}
 
 Wavelets are similar to the FT in which they deconstruct a signal using representations of other signals.
+The key difference is that wavelet signals are finite in time and frequency as opposed to sin and cos signals, which can carry on forever.
+This allows us to extract information from a signal with respect to time and location.
+
+The use of wavelets is called the wavelet transform, which is a technique in which a signal is analyzed using different versions of a dilated and translated basis functions called the mother wavelet.
+There are two types of wavelet transformations, discrete and continuous.
+In this thesis, we focus on discrete.
+Discrete Wavelet Transform uses a discrete set of wavelet scales and translations which decompose the signal into a mutually orthogonal set of wavelets.
+
+We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank, which means it is composed of cascading high-pass and low-pass filters.
+This gives us the advantage of splitting a signal into several frequency sub-bands.
+Wavelet decompositions give us the advantage of gaining features that take into account frequency and time domains.
+
+% Transformation 3: Wavelets
+
+% Wavelets are similar to the FT in which they deconstruct a signal using representations of other signals.
 The key difference is wavelet signals are finite in time and frequency as opposed to sin and cos signals which can carry on forever.
@@
 
-The use of wavelets is called the wavelet transform which is a technique in wich a signal is analyzed using different versions of a dilated and translated basis functionss called the mother wavelet.
+% The use of wavelets is called the wavelet transform which is a technique in which a signal is analyzed using different versions of a dilated and translated basis functionss called the mother wavelet.
 There are two types of wavelet transformations, discrete and continous.
@@
 
-DWT uses a discrete set of wavelet scales and translations which decompose the signal into mutually orthogional set of wavelets.
+% DWT uses a discrete set of wavelet scales and translations which decompose the signal into mutually orthogional set of wavelets.
 
-One advantage of wavelets is they have varying window sizes so they can be wide for slow frequencies and narrow for faster ones this is handled by the property called scale [add figure].
+% One advantage of wavelets is they have varying window sizes so they can be wide for slow frequencies and narrow for faster ones this is handled by the property called scale [add figure].
 
-wavelets
+% wavelets
 
-Some properties of wavelets are orthoganility, zero mean, and are efficient at representing localized data and functions.
+% Some properties of wavelets are orthogonality, zero mean, and are efficient at representing localized data and functions.
 
-Db4 wavelet http://wavelets.pybytes.com/wavelet/db4/
+% Db4 wavelet http://wavelets.pybytes.com/wavelet/db4/
 
-Model 1:
+% Model 1:
+% Model 2: 
+% Model 3:
 
-Model 2: 
-Model 3: `'
+% In the preprocessing stage we tried numerous combinations of features but found the best performing features to come from wavelet decomposition.
 
-In the preprocessing stage we tried numerous combinations of features but found the best performing features to come from wavelet decomposition.
+% Wavelets
 
-Wavelets
-
-We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank which means it’s composed of cascading high-pass and low-pass filters.
+% We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank which means it’s composed of cascading high-pass and low-pass filters.
 This gives us the advantage of splitting a signal into several frequency sub-bands.
@@
 
-[explain why they performed better]
+% [explain why they performed better]
 
-[alternatives]
+% [alternatives]
 
 \section{Data Shape}
-[Shape of data comparison]
+% [Shape of data comparison]
 
-A neural network requires all data to be the same shape but because we downsample in wavelet decomposition, each data size is halved until making it impossible to place all levels of decomposition into the same dataset.
-In order to circumvent this we tried two different methods, resampling the data, and treating each level of decomposition as its own dataset.
+A neural network requires all data to be the same shape.
+However, because we downsample in wavelet decomposition, each data size is halved until making it impossible to place all levels of decomposition into the same dataset.
+In order to circumvent this, we tried two different methods, resampling the data and treating each level of decomposition as its own dataset.
 
@@
 We use spline interpolation in order to resize each level of decomposition into a single array.
-We tested different sizes of interpolation from N to N/3 where N is the size of the largest discrete signal size and found no drop in performance.
+We tested different sizes of interpolation from N to N/3, where N is the size of the largest discrete signal size and found no reduction in performance.
 
@@
 \chapter{Models}
-[introduce models definitions]
-[EXPLAIN LOSS AND HOW ACCURACY IS MEASURED]
+% [introduce models definitions]
+% [EXPLAIN LOSS AND HOW ACCURACY IS MEASURED]
 
 \section{Definition}
-This section gives an overview of the theoretical background of concepts discussed in this work.
-
-Machine learning algorithms can simply be seen as learning a target function (f) that maps input data (X) to an output (Y).
-There are several techniques to make this work but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short Term Memory networks (LSTM).
+Machine learning algorithms can be seen as learning a target function (f) that maps input data (X) to an output (Y).
+There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short Term Memory networks (LSTM).
 Nonparametric algorithms are algorithms that attempt to make minimal assumptions about the form of the function so they can learn any form from data provided to it.
-An example would be a neural network because it has no prior knowledge about what it’s classifying and attempts to generalize any new data points that it has not seen before.
+An example would be a neural network because it has no prior knowledge about what it is classifying and attempts to generalize any new data points that it has not seen before.
 
 \subsection{CNN}
-[cnn]
+% [cnn]
 
@@
 
-Convolutional neural networks perform feature learning via non-linear transformations implemented as a series of nested layers.
-The input data is a multidimensional data array, called tensor, that is presented at the visible layer.
-This is typically a grid-like topological structure, e.g.
-time-series data, which can be seen as a 1D grid taking samples at regular time intervals, pixels in images with a 2D layout, a 3D structure of videos, etc.
-Then a series of hidden layers extract several abstract features.
-Those layers are “hidden” because their values are not given.
-Instead, the deep learning model must determine which data representations are useful for explaining the relationships in the observed data.
-Each layer consists of several kernels that perform a convolution over the input; therefore, they are also referred to as convolutional layers.
-Kernels are feature detectors, that convolve over the input and produce a transformed version of the data at the output.
-Those are banks of finite impulse response filters as seen in signal processing, just learned on a hierarchy of layers.
-The filters are usually multidimensional arrays of parameters that are learnt by the learning algorithm [24] through a training process called backpropagation.
+Convolutional neural networks perform feature learning via non-linear transformations implemented as a series of layers.
+The input data is a multidimensional array, called a tensor.
+The input data is passed through an input layer, followed by a series of hidden layers to extract features, and finally, an output layer, which in the case of classification, gives a probability for each class.
+
+Hidden layers are crucial to neural networks because it is how a model can determine which data representations are useful for explaining the relationships in the given data.
+Each layer consists of several kernels that perform a convolution over the input; therefore, they referred to as convolutional layers.
+Kernels are feature detectors that convolve over the input and produce a transformed version of the data at the output.
 
@@
 
-[lstm]
+% [lstm]
 Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture[1] used in the field of deep learning.
@@
 
-\section{black box}
-Our model is constructed of four different architectures, 2 LSTM and 2 CNN’s, that are concatenated at the dense layer.
-The four features we use are amplitude and phase of each level of decomposition for HV and VV.
-We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other.
-Wavelet decomposition produces a complex value but that is unusable in a standard neural network due to loss of information unless we move to a complex LSTM but as stated in {} there isn’t a significant increase in accuracy.
+\section{mtf}
 
-\section{white box}
-Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method.
-We then pass those images through a CNN and those that are correctly classified are shown to the user.
+We approach our MTF is slightly different than  because instead of performing it over all samples in the dataset we perform it for every signal.
 
-In order to classify those images we created a CNN for it specifically and avoided using our existing CNN in order to make it deeper and increase performance when images aren’t that classifiable.
-We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it.
+The three attributes we use are values are the absolute value of HV SBR, the absolute value of VV SBR, and a mathematical combination of the two (Figure X).
 
-\subsection{mtf}
-We approach our MTF is slightly different than [] because instead of performing it over all samples in the dataset we perform it for every signal. The three attributes we use are values o
+[add figure]
 
@@
 \section{Layout}
-
-\chapter{Problem Statement}
 
@@
 approach and consider each M to be a separate color channel of RGB where
-Mangular is red, Mtime is blue, Mradial is green. We perform linear interpolation
-on each M individually to extend the probabilities to be within the range of 0 and 255.
-
-\section{}
+Mangular is red, Mtime is blue, Mradial is green.
 
 \chapter{Main Results}
+
+\section{black box}
+Our black-box model is constructed of four different architectures, 1 LSTM and 2 CNN’s, that are concatenated at the dense layer.
+The four features we use are amplitude/phase of each level of decomposition for HV and VV.
+We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other.
+Figure {X} depicts our model.
+
+To test our model we split our data into training and test sets by randomly selecting 70\% of our data as test and left the other 30\% as our test set.
+We then performed 3 fold cross validation and took the average of our results as our final outcome.
+
+\section{10k data}
+Given that this is much larger than our other data we split our data into 10\% training and 90\% test. We found no drop in performance nor in accuracy.
+
+[add table for results]
+
+\section{white box}
+Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method.
+We then pass those images through a CNN and those that are correctly classified are shown to the user.
+
+In order to classify those images we created a CNN for it specifically and avoided using our existing CNN in order to make it deeper and increase performance when images aren’t that classifiable.
+We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it.
+
+[add results]
 

### 2020-03-14 paper/main.tex
new work

@@
    Committee Members: \\
+   Ju Cheol Moon, Ph.D \\
+   Roman Tankelevich, Ph.D \\
+   Seok-Chul Kwon, Ph.D
+
+   \vspace*{\baselineskip}
+
+   College Designee: \\
+   Hamid Rahai, Ph.D.
 
@@
 \chapter*{Abstract}
+
+% In this thesis, we study interpretable machine learning as applied to complex-valued time series data.
+Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.
+These methods, however, fall short of allowing users to visualize patterns within their dataset.
+
+% To address this issue, we propose an interpretable hybrid model that can be extended to any time series dataset.
+Our architecture is composed of two models, an ensemble method of classifiers that functions as a black-box method, and a white-box model that encodes time series as an image, attempts to classify them through a CNN, and outputs all images correctly classified in both black-box and white-box models.
+
 \addcontentsline{toc}{chapter}{Abstract}
+\chapter{Acknowledgments}
 
@@
 \mainmatter
+\chapter{Introduction}
+\label{chap:introduction}
 
-\chapter{Introduction}
+As time series datasets become ever so popular the need for deep learning models becomes greater.
+In our black-box model, we analyze deep learning models such as CNN’s which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on \cite{wang_hybrid_2019}.
+So, as black-box models get more popular the need for an interpretable accessory is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
 
-[explain what black box and white box is]
-In this thesis, we study interpretable machine learning as applied to complex-valued time series.
-Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.
-
-[add more reasons]
-[add how most time series are classified]
-These methods, however, fall short of allowing users to visualize patterns within their dataset and ().
-
-To address this issue, we propose an interpretable hybrid  model that can be extended to any time series dataset.
-Our architecture is composed of two models, an ensemble method of classifiers that functions as a black box method, and a white-box model that encodes time series as images, attemps to classify them through a CNN, and outputs all images correctly classified in both black-box and white-box method.
-
-In our black-box model, we analyze  deep learning models such as CNN’s which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on.
-So, as black-box models get more popular the need for an interpretable accessory?
-is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
-
-\section{The need for interpretable models in management problems}
-The growth of mobile devices and demand for wireless data has created a need for igh quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.
+% \section{The need for interpretable models in management problems}
+% The growth of mobile devices and demand for wireless data has created a need for igh quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.
 however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments.
-This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
+% This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
 
 \section{Contributions}
-[fill in]
+We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model.
+Our final results show how robust our black-box method is as well as support for our white-box model.
 
-\section{Data}
+\chapter{Data}
 
@@
 Spatial Modulation is a transmission technique that uses multiple antennas.
-It maps a block of information bits to two units, one is a symbol chosen from a constellation diagram, and the second one is a unique transmit antenna number that is chosen from a set of transmit antennas.
+It maps a block of information bits to two units, one is a symbol chosen from a constellation diagram, and the second one is a unique transmit antenna number that is chosen from a set of transmit antennas \cite{mesleh_spatial_2008}.
 The method in which transmit antennas send and receive radio waves can be described through polarization.
@@
 The two properties HV and VV, are given to us in raw discrete format.
-Both properties are consist of time and amplitude values but vary in length and in how they are processed.
+Both properties consist of time and amplitude values but vary in length and in how they are processed.
 The raw signal consists of over 31,000 data points with the exact time it was received.
-The size of the discrete data can vary from 22 to 44 data points, and is a processed representation of the raw signal that has equally spaced values.
+The size of the discrete data can vary from 22 to 44 data points and is a processed representation of the raw signal that has equally spaced values.
+In this thesis we focus on discrete data given that it is closer to datasets in real-world situations.
 
-[ What is raw signal and discrete, compare them, ]
-[explain hv, vv]
+% [ What is raw signal and discrete, compare them, ]
+% [explain hv, vv]
+
+\chapter{Related Work}
+\label{chap:relatedwork}
+
 \chapter{Preprocessing}
+\label{chap:preprocessing}
+
 Radio signal data is presented in a complex-valued format that is unusable in a typical neural network, so in an attempt to make our model robust and reduce any complexity, we focused on using real values as features.
@@
 
-\subsection{STFT/ FT}
+\subsection{Fourier Transform}
 % [Recall we can write a complex number in terms of its magnitude and phase
@@
 % Transformation 1 : STFT/ FT
-[what is it]
+% [what is it]
 
-[fft formula]
+% [fft formula]
 The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions.
@@
 
-One of the shortcomings of the FT is rooted in the Heisenberg Uncertainty Principle (HUP) [quote?].
+One of the shortcomings of the FT is rooted in the Heisenberg Uncertainty Principle (HUP) \cite{hill_uncertainty_nodate}.
 The HUP states that the position and velocity of an object cannot be simultaneously measured, which can be applied to the time-frequency information of a signal.
@@
 
-\subsection{Shortcoming stft}
-[stft]
-The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.[Ref] 
+\subsection{Short-Time Fourier Transform}
+% [stft]
+The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.
+\cite{hill_uncertainty_nodate}
 Computing STF requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
@@
 
+% [purpose and motivation? Behind making it ang/abs]
+
+% [definition plus method]
+
+\subsection{Sliding Window Method}
+
+We propose a method similar to STFT but to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
+This performed as well as STFT and gave us the advantage of reducing the dimensionality of our data.
+
 \subsection{Polar Coordinates}
 
-Transformation 2: Amplitude / Phase
+% Transformation 2: Amplitude / Phase
 The representation of a complex number as a sum of a real and imaginary number, z = x + iy, is called its Cartesian representation.
 For every cartesian point, we calculate its radius and angular distance which are real values and use those as features.
-
-[purpose and motivation? Behind making it ang/abs]
-
-[definition plus method]
-
-\subsection{Transformation 2: Sliding window }
-
-Transformation 2: Sliding window 
-We propose a method similar to STFT but to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
+This method outperformed the previous three methods but limited our feature size which made it difficult to get consistent accuracy across all three of our datasets.
 
@@
 Wavelets are similar to the FT in which they deconstruct a signal using representations of other signals.
-The key difference is that wavelet signals are finite in time and frequency as opposed to sin and cos signals, which can carry on forever.
+The key difference is that wavelet signals are finite in time and frequency as opposed to sin and cos signals, which can carry on forever \cite{strang_wavelets_1996}.
 This allows us to extract information from a signal with respect to time and location.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.4]{../media/wavelet_families.png}
+   \end{center}
+   \caption{Example of different types of wavelets.}
+   \label{fig:reinforcementAgent}
+\end{figure}
 
@@
 We use spline interpolation in order to resize each level of decomposition into a single array.
-We tested different sizes of interpolation from N to N/3, where N is the size of the largest discrete signal size and found no reduction in performance.
+We tested different sizes of interpolation from N to N/3, where N is the size of the largest discrete signal size, and found no reduction in performance.
 
 \subsection{Each level}
-Figure {X} depicts our model in which each level of decomposition is treated as its own dataset. We found no difference in accuracy compared to the resampling method.
+Figure \ref{fig:multimodal_overview} depicts our model in which each level of decomposition is treated as its own dataset.
+We found no difference in accuracy compared to the resampling method.
+However, given that related work we found treated each decomposition level as its own dataset, we decided to go with this method.
 
 \chapter{Models}
+\label{chap:models}
+
 % [introduce models definitions]
@@
 
-Convolutional neural networks perform feature learning via non-linear transformations implemented as a series of layers.
+Convolutional Neural Networks perform feature learning via non-linear transformations implemented as a series of layers.
 The input data is a multidimensional array, called a tensor.
 The input data is passed through an input layer, followed by a series of hidden layers to extract features, and finally, an output layer, which in the case of classification, gives a probability for each class.
+
+% create your own design
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.3]{../media/nn_highlight.jpeg}
+   \end{center}
+   \caption{Example of a Convolutional Neural Network.}
+   \label{fig:reinforcementAgent}
+\end{figure}
 
@@
 % [lstm]
-Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture[1] used in the field of deep learning.
+Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture used in the field of deep learning.
 Unlike standard feedforward neural networks, LSTM has feedback connections.
-It can not only process single data points (such as images), but also entire sequences of data (such as speech or video).
-For example, LSTM is applicable to tasks such as unsegmented, connected handwriting recognition[2], speech recognition[3][4] and anomaly detection in network traffic or IDS's (intrusion detection systems).
+It can process not only single data points but also entire sequences of data.
 
-A common LSTM unit is composed of a cell, an input gate, an output gate and a forget gate.
-The cell remembers values over arbitrary time intervals and the three gates regulate the flow of information into and out of the cell.
+A standard LSTM unit is composed of a cell, an input gate, an output gate, and a forget gate.
+The cell remembers values over arbitrary time intervals, and the three gates regulate the flow of information into and out of the cell.
 
-LSTM networks are well-suited to classifying, processing and making predictions based on time series data, since there can be lags of unknown duration between important events in a time series.
+LSTM networks are well-suited to classifying, processing, and making predictions based on time series data since there can be lags of unknown duration between significant events in a time series.
 LSTMs were developed to deal with the exploding and vanishing gradient problems that can be encountered when training traditional RNNs.
-Relative insensitivity to gap length is an advantage of LSTM over RNNs, hidden Markov models and other sequence learning methods in numerous applications.
 
-\subsection{multimodal}
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.5]{../media/lstm_highlight.png}
+   \end{center}
+   \caption{Example of a LSTM unit.}
+   \label{fig:reinforcementAgent}
+\end{figure}
 
-[multimodal]
-Due to the superior performance and computationally tractable representation capability (in vector spaces) in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks [38, 39, 44].
-Typically, domain-specific neural networks are used on different modalities to generate their representations and the individual representations are merged or aggregated.
-Finally, the prediction is made on top of aggregated representation usually with another a neural network to capture the interactions between modalities and learn complex function mapping between input and output.
-Addition (or average) and concatenation are two common aggregation methods, i.e.,
-The network structure is illustrated in Figure 1(b) .
-The arrows are function mappings or computing operations.
-The dotted boxes are representations of single and combined modality features.
-We call them additive combinations because their critical step is to add modality hidden vectors (although often in a nonlinear way).
+\subsection{Multimodal Deep Learning}
 
-[explain architectures]
+% [multimodal]
+Due to the superior performance and computationally tractable representation capability in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks \cite{kim_multimodal_2018}.
+Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated.
+Finally, the prediction is made on top of aggregated representation, usually with another neural network to capture the interactions between modalities and learn complex function mapping between input and output.
+
+% [explain architectures]
 
@@
 We attempted to make it deeper but found inconsistent performance across all of our datasets.
-[add figure]
+% [add figure]
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/cnn_model.png}
+   \end{center}
+   \caption{CNN architecture used in our black-box model.}  
+    \label{fig:cnn_model}
+\end{figure}
+
 \subsection{LSTM}
@@
 
-[add figure]
+% [add figure]
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/lstm_model.png}
+   \end{center}
+   \caption{LSTM architecture used in our black-box model.}
+   \label{fig:lstm_model}
+\end{figure}
 
-\subsection{Final method}
-We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure (X).
-The multimodal architecture allows our data to vary in size and given that all four levels of decomposition have varying length due to downsampling, we propose each model to learn the representation of each distinct feature.
-For this architecture we propose two models, a 1D CNN and an LSTM, in which we employ two CNN’s and 1 LSTM.
-They are all concatenated after their respective dense layer which performs the final   We tried numerous variations and saw no impact on performance.
-[add figure]
+\subsection{Final Model}
 
-\section{mtf}
+We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure \ref{fig:multimodal_overview}.
+The multimodal architecture allows our data to vary in size and given that all three levels of decomposition have varying length due to downsampling, we propose each model to learn the representation of each distinct feature.
+For this architecture we propose two architectures, a 1D CNN and an LSTM, in which we employ two CNN’s and 1 LSTM.
+They are all concatenated after their respective dense layer after which they perform classication via a softmax layer.
+% [add figure]
 
-We approach our MTF is slightly different than  because instead of performing it over all samples in the dataset we perform it for every signal.
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/multimodal_overview.png}
+   \end{center}
+   \caption{Overview of our black-box model.}
+   \label{fig:multimodal_overview}
+\end{figure}
 
-The three attributes we use are values are the absolute value of HV SBR, the absolute value of VV SBR, and a mathematical combination of the two (Figure X).
+% \section{mtf}
 
-[add figure]
+% We approach our MTF is slightly different than  because instead of performing it over all samples in the dataset we perform it for every signal.
 
-\section{Notation}
+% The three attributes we use are values are the absolute value of HV SBR, the absolute value of VV SBR, and a mathematical combination of the two (Figure X).
 
-\section{Layout}
-
-\chapter{Related Work}
+% [add figure]
 
 \chapter{Markov Transition Fields}
-We propose a framework similar to [] for encoding dynamical transition statistics,
-but we continue their work by considering ith order Markov transition probabilities.
-Given a time series X, we decompose its magnitude axis into two separate
-properties by representing it as a polar coordinate, Xangular and Xradial. We then
-identify the Q quantile bins for both properties and the temporal values Xtime
-and assign each x in Xangular, Xradial, Xtime to its corresponding bin qj (j in
-[1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
-Wradial, Wtime, by counting transitions among quantile bins in the manner of a
-ith order Markov chain along the time axis.
+\label{chap:mtf}
 
-W does not take into account the temporal axis so to prevent any information
-loss we construct a Markov Transition Field (MTF) for each W. The MTF
-denotes the probability of transitioning from qi to qj for each x in X. This,
-in turn, allows us to consider the transition probability on the magnitude and
-temporal axis.
+% We propose a framework similar to [] for encoding dynamical transition statistics,
+% but we continue their work by considering ith order Markov transition probabilities.
+% Given a time series X, we decompose its magnitude axis into two separate
+% properties by representing it as a polar coordinate, Xangular and Xradial. We then
+% identify the Q quantile bins for both properties and the temporal values Xtime
+% and assign each x in Xangular, Xradial, Xtime to its corresponding bin qj (j in
+% [1, Q]). Thus we construct three QxQ adjacency transition matrices, Wangular,
+% Wradial, Wtime, by counting transitions among quantile bins in the manner of a
+% ith order Markov chain along the time axis.
 
-As described in [] the MTF encodes the multi-span transition probabilities
-of the time series, but given that we have three different M’s we modify their
-approach and consider each M to be a separate color channel of RGB where
-Mangular is red, Mtime is blue, Mradial is green.
+% W does not take into account the temporal axis so to prevent any information
+% loss we construct a Markov Transition Field (MTF) for each W. The MTF
+% denotes the probability of transitioning from qi to qj for each x in X. This,
+% in turn, allows us to consider the transition probability on the magnitude and
+% temporal axis.
+
+% As described in [] the MTF encodes the multi-span transition probabilities
+% of the time series, but given that we have three different M’s we modify their
+% approach and consider each M to be a separate color channel of RGB where
+% Mangular is red, Mtime is blue, Mradial is green.
+
+% \section{Markov Transition Field}
+
+We propose a framework similar to \cite{wang_encoding_nodate}  for encoding dynamical transition statistics, but we continue their work by i\textsuperscript{th} order Markov transition probabilities.
+
+Given a time series X, we decompose its magnitude axis into three separate properties, P\textsubscript{1},  P\textsubscript{2},  P\textsubscript{3}.
+We then identify the Q quantile bins for each property  and assign each $x \in P\textsubscript{1},  P\textsubscript{2},  P\textsubscript{3}$ to its corresponding bin q\textsubscript{j} (j in [1, Q]).
+Thus, we construct three QxQ adjacency transition matrices, W\textsubscript{1}, W\textsubscript{2}, W\textsubscript{3}, by counting transitions among quantile bins in the manner of a i\textsuperscript{th} order Markov chain along the time axis.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/mtf_q.png}
+   \end{center}
+   \caption{Example of QxQ quantile bin matrix used to calculate MTF.}
+   \label{fig:mtf_q}
+\end{figure}
+
+%summation %.
+This effectively removes any sensitivity to the X’s distribution and temporal dependency to each time step t i.
+To account for any information loss after removing the temporal dependency, we follow the same approach as \cite{wang} and construct a Markov Transition Field for each property like so: %some figure%
+
+W does not take into account the temporal axis so to prevent any information loss we construct a Markov Transition Field, M, for each W.
+The MTF denotes the probability of transitioning from q\textsubscript{i} to q\textsubscript{j} for each $x \in P$.
+This, in turn, allows us to consider the transition probability on the magnitude and temporal axis.
+
+\begin{figure}[h]
+  \centering
+  M =
+  \[\begin{bmatrix}
+
+  \text{W\textsubscript{ij}} | \text{$x\textsubscript{1} \in q\textsubscript{i}$, $x_{1} \in q\textsubscript{j}$}
+  & \dots &
+
+  \text{W\textsubscript{ij}}  |  \text{$x_{1} \in q\textsubscript{i}$, $x_{n} \in q\textsubscript{j}$}
+
+  \\
+
+  \vdots &  \ddots & \\
+
+  \text{W\textsubscript{ij}} | \text{$x_{n} \in q\textsubscript{i}$, $x_{1} \in q\textsubscript{j}$}
+  & \dots &
+  \text{W\textsubscript{ij}} | \text{$x_{n} \in q\textsubscript{i}$, $x_{n} \in q\textsubscript{j}$} \\
+
+  \end{bmatrix}\]
+  \caption{Markov Transition Field where  W\textsubscript{ij} denotes the transition probability from quantile i to j.}
+  \label{fig:mtf}
+\end{figure}
+
+As described in \cite{wang_encoding_nodate} the MTF encodes the multi-span transition probabilities of the time series, but given that we have three different M's we modify their approach and consider each M to be a separate color channel of RGB where M\textsubscript{1} is red, M\textsubscript{2} is blue, M\textsubscript{3} is green.
+Since each row in M is a probability from 0 to 1 we multiply it by 255 to get a color value.
 
 \chapter{Main Results}
+\label{chap:results}
 
-\section{black box}
-Our black-box model is constructed of four different architectures, 1 LSTM and 2 CNN’s, that are concatenated at the dense layer.
-The four features we use are amplitude/phase of each level of decomposition for HV and VV.
+\section{Black-Box Model}
+Our black-box model is constructed of three different architectures, 1 LSTM and 2 CNN’s, that are concatenated at the dense layer.
+The three features we use are amplitude/phase of each level of decomposition for HV and VV.
 We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other.
-Figure {X} depicts our model.
 
@@
 
-\section{10k data}
-Given that this is much larger than our other data we split our data into 10\% training and 90\% test. We found no drop in performance nor in accuracy.
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/results.png}
+   \end{center}
+   \caption{Results}
+   \label{fig:final_results}
+\end{figure}
 
-[add table for results]
+\section{Dataset B}
 
-\section{white box}
+Given that this dataset is ten times larget than our other two datasets we wanted to try different training and test splits, namely we went as low as 8\% for training data and used the remaining 92\% as test data.
+We ran the same number of cross validation as our previous tests and found minimal difference in performance.
+
+% [add table for results]
+
+\section{White-Box Model}
 Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method.
@@
 
-[add results]
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=.7]{../media/alexnet.png}
+   \end{center}
+   \caption{CNN architecture for classifying images generated through MTF's.}
+   \label{fig:mtf_cnn_results}
+\end{figure}
+
+% [add results]
 
@@
 \label{chap:conclusion}
+Throughout this thesis, we have stressed the importance of a hybrid machine learning algorithm as well as providing a framework that can successfully classify and visualize different classes when applied to time-series data.
 
-% \bibliography{../references/main}
-% \bibliographystyle{plain}
+We proposed an extension to existing work described in [] so users can visualize their data while being able to classify them with 60\% accuracy on our dataset and model.
+We also created a robust black-box model consisting of two different deep learning architectures that consistently provided competitive accuracy.
+
+In the future, we would like to explore other methods for our white-box model that would act as an alternative to Markov Transition Fields in hopes that it produces higher accuracy than our current model.
+Additionally, we would like to extensively test our model on different time-series datasets, not just radio signal datasets.
+
+\bibliography{../references/main}
+\bibliographystyle{plain}
 

### 2020-03-14 paper/main.tex
updated intro

@@
 
+In this thesis, we study interpretable machine learning as applied to complex-valued time-series data.
+Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.
+These methods, however, fall short of allowing users to visualize patterns within their dataset.
+
+To address this issue, we propose an interpretable hybrid model that can be extended to any time series dataset.
+In the majority of existing work in the field of interpretability, black-box models tend to outperform white-box models consistently.
+However, instead of relying purely on one method, we propose a collaboration between the two.
+This gives us the performance of a black-box while allowing us to visualize features of the majority of our dataset.
+
+To accomplish this, we created a framework composed of two models, an ensemble method of classifiers that functions as a black-box method, and a white-box model that encodes time series as images.
+The white-box model attempts to classify them through a neural network and outputs all images correctly classified in both black-box and white-box models.
+
 \addcontentsline{toc}{chapter}{Abstract}
@@
 As time series datasets become ever so popular the need for deep learning models becomes greater.
-In our black-box model, we analyze deep learning models such as CNN’s which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on \cite{wang_hybrid_2019}.
-So, as black-box models get more popular the need for an interpretable accessory is needed so we propose a hybrid model consisting of a black box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
+In our black-box model, we analyze deep learning models such as Convolutional Neural Networks which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on inside \cite{wang_hybrid_2019}.
+So, as black-box models get more popular and increase in complexity the need for an interpretable accessory is needed so we propose a hybrid model consisting of a black-box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
+
+The goal of our work is not to find an alternative to black-box methods but to provide a way for users to have an understanding of what features are prevalent in their dataset.
 
@@
 We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank, which means it is composed of cascading high-pass and low-pass filters.
-This gives us the advantage of splitting a signal into several frequency sub-bands.
-Wavelet decompositions give us the advantage of gaining features that take into account frequency and time domains.
+This allows us to split a signal into several frequency sub-bands.
+We focus on wavelet decomposition as our feature extractor because it outperformed all other methods in terms of accuracy and training time.
 

### 2020-03-15 code/main.py
new figures

@@ -1,0 +1,212 @@
+import numpy as np
+from scipy.io import loadmat
+import matplotlib.pyplot as plt
+import scipy.ndimage
+
+def resize_(before, after_size=5000):
+    before_size = before.shape[0]
+    ratio = after_size / before_size
+    return scipy.ndimage.zoom(before, ratio)
+
+
+import os, glob
+from scipy import signal
+from pywt import wavedec
+import pywt
+
+
+def generateData(path):
+    max_lev = 3
+
+    class_names = ['h_HV_SBR_1_discrete', 'h_HV_SBR_2_discrete', 'h_HV_SBR_3_discrete']
+    class_names_vv = ['h_VV_SBR_1_discrete', 'h_VV_SBR_2_discrete', 'h_VV_SBR_3_discrete']
+
+    data1=[]
+    data2=[]
+    data3 = []
+    num_classes = 3
+
+    c= 0
+    for _file in glob.glob(path):
+        filename = _file
+        file_contents = loadmat(filename)
+
+        c+=1
+        if c == 11:
+            break
+
+        for idx in range(num_classes):
+            fc = file_contents[class_names[idx]]
+            fc_vv = file_contents[class_names_vv[idx]]
+            
+            #normalize data
+            data = fc / np.linalg.norm(fc)
+            data_vv = fc_vv /np.linalg.norm(fc_vv)
+            
+            #Assign data to a variable so we can overwrite it after every level of decomposition
+            hv = data[:22]
+            vv = data_vv[:22]
+
+            for level in range(0, max_lev):
+
+                cA_hv, cD_hv = pywt.dwt(hv, 'db4', axis=0)
+                cA1_vv, cD_vv = pywt.dwt(vv, 'db4', axis=0)
+
+                hv = cD_hv
+                vv = cD_vv
+
+                if level == 0:
+                    data1.append(np.abs(cA_hv))
+
+                if level == 1:
+                    data2.append(np.abs(cA_hv))
+
+                if level == 2:
+                    data3.append(np.abs(cA_hv))  
+
+    data1 = np.asarray(data1)
+    data2 = np.asarray(data2)
+    data3 = np.asarray(data3)
+
+    train_test_split = .7
+
+    return processData(data1, data2, data3, train_test_split)
+
+
+'''
+    Function: processData
+    Purpose: Returns training and test data with the option of having them randomized
+'''
+def processData(data1, data2, data3, train_split = .7, random=True):
[…]

### 2020-03-16 paper/main.tex
new figures
formatting

@@
 % To address this issue, we propose an interpretable hybrid model that can be extended to any time series dataset.
-Our architecture is composed of two models, an ensemble method of classifiers that functions as a black-box method, and a white-box model that encodes time series as an image, attempts to classify them through a CNN, and outputs all images correctly classified in both black-box and white-box models.
+Our architecture is composed of two models, an ensemble method of classifiers that functions as a black-box model, and a white-box model that encodes time series as an image, attempts to classify them through a CNN, and outputs all images correctly classified in both black-box and white-box models.
 
@@
 However, instead of relying purely on one method, we propose a collaboration between the two.
-This gives us the performance of a black-box while allowing us to visualize features of the majority of our dataset.
+This allows for the performance of a black-box model while providing a more precise visualization of our dataset.
 
-To accomplish this, we created a framework composed of two models, an ensemble method of classifiers that functions as a black-box method, and a white-box model that encodes time series as images.
+To accomplish this, we created a framework composed of two models, an ensemble method of classifiers that functions as a black-box model, and a white-box model that encodes time series as images.
 The white-box model attempts to classify them through a neural network and outputs all images correctly classified in both black-box and white-box models.
@@
 \chapter{Acknowledgments}
-
+[to do]
 \tableofcontents
@@
 
-As time series datasets become ever so popular the need for deep learning models becomes greater.
-In our black-box model, we analyze deep learning models such as Convolutional Neural Networks which have become increasingly popular and widely used in the last decade but their complex nature makes it difficult for users to understand what is going on inside \cite{wang_hybrid_2019}.
-So, as black-box models get more popular and increase in complexity the need for an interpretable accessory is needed so we propose a hybrid model consisting of a black-box for classification and an interpretable white-box model which highlights characteristics of a time series for users to understand.
+As time-series datasets become ever more popular, the need for deep learning models becomes that much greater.
+In our black-box model, we analyze deep learning models such as Convolutional Neural Networks, which have become increasingly popular and widely used in the last decade.
+However, their complex nature results in a lack of understanding of its inner workings \cite{wang_hybrid_2019}.
+As the use and complexity of black-box models rises, so too does the need for an interpretable accessory.
+For this reason, we propose a hybrid model consisting of a black-box for classification and an interpretable white-box which highlights characteristics of a time series for users to understand.
 
-The goal of our work is not to find an alternative to black-box methods but to provide a way for users to have an understanding of what features are prevalent in their dataset.
+The goal of our work is not to find an alternative to black-box models, but to provide a way for users to have an understanding of what features are prevalent in their dataset.
 
@@
 We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model.
-Our final results show how robust our black-box method is as well as support for our white-box model.
+Our final results show how robust our black-box model is with its competitive accuracy.
+Additionally, we show the effectiveness of our white-box model in understanding the features of a dataset not normally exposed through a black-box model.
 
@@
 We analyzed our model on three radio signal datasets generated using Spatial Modulation.
-Spatial Modulation is a transmission technique that uses multiple antennas.
-It maps a block of information bits to two units, one is a symbol chosen from a constellation diagram, and the second one is a unique transmit antenna number that is chosen from a set of transmit antennas \cite{mesleh_spatial_2008}.
-The method in which transmit antennas send and receive radio waves can be described through polarization.
+Spatial Modulation is a transmission technique that uses multiple antennae which maps a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}.
+The method in which transmit antennae send and receive radio waves can be described through polarization.
 Two popular basis polarizations are horizontal linear, H, and vertical linear, V.
-Our datasets contain data horizontally transmitted and vertically received (HV) and data vertically transmitted and vertically received.
+Our datasets contain data horizontally transmitted and vertically received (HV), as well as data vertically transmitted and vertically received (VV).
 
-The two properties HV and VV, are given to us in raw discrete format.
+The two properties, HV and VV, are given to us in a raw and discrete format.
 Both properties consist of time and amplitude values but vary in length and in how they are processed.
@@
 The size of the discrete data can vary from 22 to 44 data points and is a processed representation of the raw signal that has equally spaced values.
-In this thesis we focus on discrete data given that it is closer to datasets in real-world situations.
+In this thesis, we focus on discrete data given that it is closer to datasets in real-world situations.
 
@@
 
-Radio signal data is presented in a complex-valued format that is unusable in a typical neural network, so in an attempt to make our model robust and reduce any complexity, we focused on using real values as features.
+The radio signal data is presented in a complex-valued format that is unusable in a typical neural network, so in an attempt to make our model robust and reduce any complexity, we focused on using real values as features.
 
-We tried numerous methods to extract real-valued features, including Fourier Transform, Short Time Fourier Transform, a custom sliding window method, and polar coordinates taken directly from the signals amplitude/phase value.
+We tried numerous methods to extract real-valued features, including Fourier Transform, Short-Time Fourier Transform, a custom sliding window method, and polar coordinates taken directly from the signal's amplitude/phase value.
 Before applying any of these methods, we normalized all data using L2 normalization.
@@
 The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions.
-Each basis function is a complex value of a frequency, so it allows us to view our data in the frequency domain as opposed to amplitude.
+Each basis function is a complex value of a frequency, allowing us to view our data in the frequency domain as opposed to the amplitude domain.
 
@@
 % [stft]
-The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time.
-\cite{hill_uncertainty_nodate}
-Computing STF requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
+The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time \cite{hill_uncertainty_nodate}.
+Computing STFT requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
 This allows us to view the Fourier spectrum at a more granular level, which could potentially reveal different patterns amongst signals of different classes.
 
-One of the issues with this method is that we can create very narrow window sizes, which gives us a better understanding of the data with respect to time, but we lose understanding of the frequency domain.
-Additionally, selecting an appropriate window size for segmenting the signal can be an arduous task that would require fine-tuning as well as increasing the dimensionality of our dataset.
+One of the issues with this method is that we can create very narrow window sizes, giving us a better understanding of the data with respect to time but losing an understanding of the frequency domain as a whole.
+Additionally, selecting an appropriate window size for segmenting the signal can be an arduous task that would require fine-tuning as well as increasing the dimensionality of a dataset.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/stft.png}
+   \end{center}
+   \caption{Example of Short-Time Fourier Transform.}
+   \label{fig:stft_example}
+\end{figure}
 
@@
 
-We propose a method similar to STFT but to reduce the size of the data we take the mean at each window which would turn our size into size N where N is equal to the number of windows we use.
-This performed as well as STFT and gave us the advantage of reducing the dimensionality of our data.
+Similar to STFT, we segment our signal into N windows of equal length and perform FT on every segmentation.
+However, instead of keeping the output from FT, we only keep the mean value of each segmentation.
+This performed as well as STFT and gave us the advantage of reducing the dimensionality of our dataset.
 
@@
 The representation of a complex number as a sum of a real and imaginary number, z = x + iy, is called its Cartesian representation.
-For every cartesian point, we calculate its radius and angular distance which are real values and use those as features.
-This method outperformed the previous three methods but limited our feature size which made it difficult to get consistent accuracy across all three of our datasets.
+For every cartesian point, we calculate its radius and angular distance as real values and use those as features.
+This method outperformed the previous three methods but limited our feature size, which made it difficult to get consistent accuracy across all three of our datasets.
 
@@
 
-Wavelets are similar to the FT in which they deconstruct a signal using representations of other signals.
-The key difference is that wavelet signals are finite in time and frequency as opposed to sin and cos signals, which can carry on forever \cite{strang_wavelets_1996}.
+Wavelet transforms are similar to the FT in that they deconstruct a signal using representations of other signals.
+The key difference is that wavelet signals are finite in time and frequency as opposed to sine and cosine signals, which can carry on indefinitely \cite{strang_wavelets_1996}.
 This allows us to extract information from a signal with respect to time and location.
@@
    \end{center}
-   \caption{Example of different types of wavelets.}
+   \caption{Example of different types of wavelets  \cite{strang_wavelets_1996}.}
    \label{fig:reinforcementAgent}
@@
 
-The use of wavelets is called the wavelet transform, which is a technique in which a signal is analyzed using different versions of a dilated and translated basis functions called the mother wavelet.
+A signal is analyzed using different versions of dilated and translated basis functions called the mother wavelet  \cite{strang_wavelets_1996}.
 There are two types of wavelet transformations, discrete and continuous.
-In this thesis, we focus on discrete.
-Discrete Wavelet Transform uses a discrete set of wavelet scales and translations which decompose the signal into a mutually orthogonal set of wavelets.
+In this thesis, we focus on discrete wavelet transformations, which use a discrete set of wavelet scales and translations to decompose the signal into a mutually orthogonal set of wavelets.
 
-We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank, which means it is composed of cascading high-pass and low-pass filters.
-This allows us to split a signal into several frequency sub-bands.
+We take advantage of wavelets by applying discrete wavelet transforms as a filter-bank, meaning they are composed of cascading high-pass and low-pass filters.
+This allows us to split a signal into several frequency sub-bands  \cite{strang_wavelets_1996}.
 We focus on wavelet decomposition as our feature extractor because it outperformed all other methods in terms of accuracy and training time.
@@
 A neural network requires all data to be the same shape.
-However, because we downsample in wavelet decomposition, each data size is halved until making it impossible to place all levels of decomposition into the same dataset.
-In order to circumvent this, we tried two different methods, resampling the data and treating each level of decomposition as its own dataset.
+However, because we downsample in wavelet decomposition, our data size is halved, making it impossible to place all levels of decomposition into the same dataset.
+In order to circumvent this, we tested two different methods: resampling the data and treating each level of decomposition as a unique dataset.
 
@@
 
-We use spline interpolation in order to resize each level of decomposition into a single array.
+We use spline interpolation to resize each level of decomposition into a single array.
 We tested different sizes of interpolation from N to N/3, where N is the size of the largest discrete signal size, and found no reduction in performance.
 
-\subsection{Each level}
-Figure \ref{fig:multimodal_overview} depicts our model in which each level of decomposition is treated as its own dataset.
-We found no difference in accuracy compared to the resampling method.
-However, given that related work we found treated each decomposition level as its own dataset, we decided to go with this method.
+\subsection{Each Level of Decomposition as a Dataset}
+Figure \ref{fig:multimodal_overview} depicts our model in which each level of decomposition is a unique dataset.
+This method and the resampling method both achieved similar results, so we chose to move forward with this method.
+[make new figure]
 
@@
 Machine learning algorithms can be seen as learning a target function (f) that maps input data (X) to an output (Y).
-There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short Term Memory networks (LSTM).
-Nonparametric algorithms are algorithms that attempt to make minimal assumptions about the form of the function so they can learn any form from data provided to it.
-An example would be a neural network because it has no prior knowledge about what it is classifying and attempts to generalize any new data points that it has not seen before.
+There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short-Term Memory networks (LSTM).
+Nonparametric algorithms attempt to make minimal assumptions about the form of the function to learn any form from data provided to it.
+An example would be a neural network as it has no prior knowledge of what it is classifying and attempts to generalize any new data points.
 
@@
 A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis.
-It consists of an input and output layer as well as hidden layers.
-A convolution is an operation between a vector of weights w against an input x.
-It consists of taking the dot product between m and x were in steps of a filter size defined by n.
+As seen in Figure X, it consists of an input and output layer as well as several hidden layers.
+A convolution is an operation between a vector of weights w against an input x \ref{}.
+It consists of taking the dot product between m and x in steps of a filter size defined by n.
 
@@
 The input data is a multidimensional array, called a tensor.
-The input data is passed through an input layer, followed by a series of hidden layers to extract features, and finally, an output layer, which in the case of classification, gives a probability for each class.
+The tensor is passed through an input layer, followed by a series of hidden layers to extract features, and an output layer, which in the case of classification, gives a probability for each class.
 
@@
 
-Hidden layers are crucial to neural networks because it is how a model can determine which data representations are useful for explaining the relationships in the given data.
-Each layer consists of several kernels that perform a convolution over the input; therefore, they referred to as convolutional layers.
-Kernels are feature detectors that convolve over the input and produce a transformed version of the data at the output.
+Hidden layers are crucial to neural networks because they help in determining which data representations are useful for explaining the relationships in the given data.
+Each layer consists of several kernels,  which are feature detectors that convolve over the input data and output a transformed version of it.
 
@@
 Long short-term memory (LSTM) is an artificial recurrent neural network (RNN) architecture used in the field of deep learning.
-Unlike standard feedforward neural networks, LSTM has feedback connections.
-It can process not only single data points but also entire sequences of data.
+Unlike standard feedforward neural networks, LSTM has feedback connections, processing not only single data points but also entire sequences of data.
 
 A standard LSTM unit is composed of a cell, an input gate, an output gate, and a forget gate.
-The cell remembers values over arbitrary time intervals, and the three gates regulate the flow of information into and out of the cell.
+The cell remembers values over arbitrary time intervals, and the three gates regulate the flow of information into and out of the cell \ref{}.
 
 LSTM networks are well-suited to classifying, processing, and making predictions based on time series data since there can be lags of unknown duration between significant events in a time series.
-LSTMs were developed to deal with the exploding and vanishing gradient problems that can be encountered when training traditional RNNs.
+LSTMs are developed to deal with the vanishing gradient problems that can are encountered when training traditional RNNs \ref{}.
 
@@
 Due to the superior performance and computationally tractable representation capability in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks \cite{kim_multimodal_2018}.
-Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated.
-Finally, the prediction is made on top of aggregated representation, usually with another neural network to capture the interactions between modalities and learn complex function mapping between input and output.
+Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated \ref{}.
+Finally, the prediction is made on top of aggregated representation, usually with another neural network to capture the interactions between models and learn complex function mapping between input and output.
 
@@
 
-Our CNN architecture consists of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to two dense layers (Figure X).
+Our CNN architecture consists of 3 convolutional layers each followed by a batch normalization and dropout layer then finally connected to a dense layer (Figure \ref{fig:cnn_model}).
 We attempted to make it deeper but found inconsistent performance across all of our datasets.
@@
 \subsection{LSTM}
-We wanted to keep our LSTM network as small as possible, for X purposes, so we went with two LSTM layers of 128 hidden states and a dropout rate of 4/10  followed by a dense layer of 128 connections.
-We attempted different state sizes but found 128 to be the smallest number with the best consistent performance.
+We wanted to keep our LSTM network as small as possible for simplicity, so we went with two LSTM layers of 128 hidden states and a dropout rate of 0.4  followed by a dense layer of 128 connections.
+We attempted different state sizes but found 128 to be the smallest number of units with the highest and most consistent rate of accuracy.
 
@@
 
-We propose a multimodal learning architecture for the purpose of time series classification which is illustrated in Figure \ref{fig:multimodal_overview}.
-The multimodal architecture allows our data to vary in size and given that all three levels of decomposition have varying length due to downsampling, we propose each model to learn the representation of each distinct feature.
-For this architecture we propose two architectures, a 1D CNN and an LSTM, in which we employ two CNN’s and 1 LSTM.
-They are all concatenated after their respective dense layer after which they perform classication via a softmax layer.
+We propose a multimodal learning architecture for time series classification, which is illustrated in Figure \ref{fig:multimodal_overview}.
+The multimodal architecture allows our data to vary in size, and given that all three levels of decomposition have varying lengths due to downsampling, we propose that each model learn the representation of each distinct feature.
+For this model, we propose two architectures: a 1D CNN and an LSTM.
+We employ two CNN’s and one LSTM, with each concatenated following their respective dense layer, after which classification is performed via a softmax layer.
 % [add figure]
@@
 % \section{Markov Transition Field}
+\section{Method}
 
-We propose a framework similar to \cite{wang_encoding_nodate}  for encoding dynamical transition statistics, but we continue their work by i\textsuperscript{th} order Markov transition probabilities.
+We propose a framework similar to \cite{wang_encoding_nodate}  for encoding dynamical transition statistics, but we continue their work by adding i\textsuperscript{th} order Markov transition probabilities.
 
@@
 We then identify the Q quantile bins for each property  and assign each $x \in P\textsubscript{1},  P\textsubscript{2},  P\textsubscript{3}$ to its corresponding bin q\textsubscript{j} (j in [1, Q]).
-Thus, we construct three QxQ adjacency transition matrices, W\textsubscript{1}, W\textsubscript{2}, W\textsubscript{3}, by counting transitions among quantile bins in the manner of a i\textsuperscript{th} order Markov chain along the time axis.
+Thus, we construct three QxQ adjacency transition matrices, W\textsubscript{1}, W\textsubscript{2}, W\textsubscript{3}, by counting transitions among quantile bins in the manner of an i\textsuperscript{th} order Markov chain along the time axis.
 
@@
 
-W does not take into account the temporal axis so to prevent any information loss we construct a Markov Transition Field, M, for each W.
+W does not take into account the temporal axis, so to prevent any information loss, we construct a Markov Transition Field, M, for each W.
 The MTF denotes the probability of transitioning from q\textsubscript{i} to q\textsubscript{j} for each $x \in P$.
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
   \centering
@@
 
-As described in \cite{wang_encoding_nodate} the MTF encodes the multi-span transition probabilities of the time series, but given that we have three different M's we modify their approach and consider each M to be a separate color channel of RGB where M\textsubscript{1} is red, M\textsubscript{2} is blue, M\textsubscript{3} is green.
-Since each row in M is a probability from 0 to 1 we multiply it by 255 to get a color value.
+As described in \cite{wang_encoding_nodate} the MTF encodes the multi-span transition probabilities of the time series, but given that we have three different M's, we modify their approach and consider each M to be a separate color channel of RGB where M\textsubscript{1} is red, M\textsubscript{2} is blue, M\textsubscript{3} is green.
+Since each row in M is a probability from 0 to 1, we multiply it by 255 to get a color value.
+
+\section{Classification}
+We apply MTF to our dataset and classify our images using a network similar to AlexNet, as shown in Figure \ref{fig:mtf_cnn_model}.
+We avoided using the CNN used in our black-box model in order to create a deeper and higher performing network.
+For this model, we define our input as images with dimensions 224x224, use a kernel size of 3x3, and apply max-pooling after every convolutional layer.
+We have five convolutional layers using 11, 256, 256, 384, and 384 filters, respectively, followed by two dense layers and a softmax layer.
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=.7]{../media/2d_cnn.png}
+   \end{center}
+   \caption{CNN architecture for classifying images generated through MTF's.}
+   \label{fig:mtf_cnn_model}
+\end{figure}
+
+% [add results]
 
@@
 \section{Black-Box Model}
-Our black-box model is constructed of three different architectures, 1 LSTM and 2 CNN’s, that are concatenated at the dense layer.
-The three features we use are amplitude/phase of each level of decomposition for HV and VV.
-We chose four because we wanted to take into consideration both HV / VV in the event that one had more descriptive features than the other.
+Our black-box model is constructed of three different architectures: 1 LSTM and 2 CNN’s that are all concatenated at their respective dense layer.
+The three features we use are the amplitude/phase of each level of decomposition for HV.
+We chose three because we wanted to increase the dimensionality of our data in the event that one had more descriptive features than the others.
 
-To test our model we split our data into training and test sets by randomly selecting 70\% of our data as test and left the other 30\% as our test set.
-We then performed 3 fold cross validation and took the average of our results as our final outcome.
+To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set.
+We then performed three-fold cross-validation and took the average of our results as the outcome.
 
@@
 
-Given that this dataset is ten times larget than our other two datasets we wanted to try different training and test splits, namely we went as low as 8\% for training data and used the remaining 92\% as test data.
-We ran the same number of cross validation as our previous tests and found minimal difference in performance.
+Given that this dataset is ten times larget than our other two datasets, we wanted to try different training and test splits, going as low as 8\% for training data and using the remaining 92\% as test data.
+We ran the same number of cross-validations as our previous tests and found minimal difference in performance.
 
@@
 
-\section{White-Box Model}
-Our white box is composed of a model that takes in raw data as an input and converts each signal into an image through the MTF method.
-We then pass those images through a CNN and those that are correctly classified are shown to the user.
-
-In order to classify those images we created a CNN for it specifically and avoided using our existing CNN in order to make it deeper and increase performance when images aren’t that classifiable.
-We explored several architectures, such as, ResNet, Inception-v4, and AlexNet but we decided to create an architecture similar to AlexNet due to the variation and size of datasets we used we couldn’t create a model deeper or more complex than it otherwise performance would drop because we didn’t have enough data to properly train it.
-
-\begin{figure}[h!]
-   \begin{center}
-      \includegraphics[scale=.7]{../media/alexnet.png}
-   \end{center}
-   \caption{CNN architecture for classifying images generated through MTF's.}
-   \label{fig:mtf_cnn_results}
-\end{figure}
-
-% [add results]
+\section{ MTF }
 
@@
 \label{chap:conclusion}
-Throughout this thesis, we have stressed the importance of a hybrid machine learning algorithm as well as providing a framework that can successfully classify and visualize different classes when applied to time-series data.
+Throughout this thesis, we have stressed the importance of a hybrid machine learning algorithm as well as provided a framework that can successfully classify and visualize different classes when applied to time-series data.
 
-We proposed an extension to existing work described in [] so users can visualize their data while being able to classify them with 60\% accuracy on our dataset and model.
+We proposed an extension to existing work described in \cite{wang_encoding_nodate} to help users visualize their data while performing classification  with 60\% accuracy on that dataset and model.
 We also created a robust black-box model consisting of two different deep learning architectures that consistently provided competitive accuracy.
@@
 In the future, we would like to explore other methods for our white-box model that would act as an alternative to Markov Transition Fields in hopes that it produces higher accuracy than our current model.
-Additionally, we would like to extensively test our model on different time-series datasets, not just radio signal datasets.
+Additionally, we would like to extensively test our model on different time-series datasets outside of radio signal datasets.
 

### 2020-03-16 paper/main.tex
edits

@@
 \chapter{Acknowledgments}
-[to do]
+
+I want to thank Dr. Moon for his guidance and for allowing me to work on a project this extensive. I would also like to thank my friends and family for their continuous support throughout this thesis.
+
 \tableofcontents
@@
 
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.5]{../media/overview_model.png}
+   \end{center}
+   \caption{Overview of our proposed hybrid model.}
+   \label{fig:overview_model}
+\end{figure}
+
 % \section{The need for interpretable models in management problems}
@@
 
-We analyzed our model on three radio signal datasets generated using Spatial Modulation.
-Spatial Modulation is a transmission technique that uses multiple antennae which maps a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}.
+We analyzed our model on three radio signal datasets, which we will refer to as dataset A, B, and C throughout this thesis.
+These datasets were generated through Spatial Modulation which is a transmission technique that uses multiple antennae which maps a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}.
 The method in which transmit antennae send and receive radio waves can be described through polarization.
@@
 Our datasets contain data horizontally transmitted and vertically received (HV), as well as data vertically transmitted and vertically received (VV).
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.5]{../media/dataset_info.png}
+   \end{center}
+   \caption{Size of each dataset.}
+   \label{fig:stft_example}
+\end{figure}
 
@@
 \label{chap:relatedwork}
+
+The study of interpretable hybrid models has been worked on extensively in the past few years as deep learning has become increasingly popular.
+Scientists take different approaches as to what makes a hybrid model interpretable and apply it to different domains, such as computer vision or time series.
+For example, in \cite{wang_multilevel_2018}, they focus on creating a black-box model for the classification of time-series that outperforms similar architectures.
+However, their white-box model focuses on analyzing the importance of their model's layers.
+This helps users identify the importance of wavelet decomposition but not necessarily understand their dataset.
+
+Additionally, other work we found tends to focus on non-time-series data or creating a general framework defining the trade-off between interpretability and black-box methods.
+For instance, \cite{wang_hybrid_2019} defines a set of rules to achieve a high level of classification with interpretability but does not give any competitive methods to achieve this.
+
+In order to fill the gap for a high performing interpretable model for time-series classification, we propose our hybrid model, which focuses on extensibility and robustness.
+We achieve this through a combination of deep learning models for classification as well as a unique white-box method to explaining features within classes of a given dataset.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
    \begin{center}
@@
 This method and the resampling method both achieved similar results, so we chose to move forward with this method.
-[make new figure]
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.35]{../media/wavelet_decomp.png}
+   \end{center}
+   \caption{Example of a wavelet decomposition network.}
+   \label{fig:wavelet_decomp}
+\end{figure}
 
@@
 As seen in Figure X, it consists of an input and output layer as well as several hidden layers.
-A convolution is an operation between a vector of weights w against an input x \ref{}.
+A convolution is an operation between a vector of weights w against an input x \cite{yamashita_convolutional_2018}.
 It consists of taking the dot product between m and x in steps of a filter size defined by n.
@@
    \end{center}
-   \caption{Example of a Convolutional Neural Network.}
-   \label{fig:reinforcementAgent}
+   \caption{Example of a Convolutional Neural Network \cite{yamashita_convolutional_2018}.}
+   \label{fig:cnn_example}
 \end{figure}
@@
 A standard LSTM unit is composed of a cell, an input gate, an output gate, and a forget gate.
-The cell remembers values over arbitrary time intervals, and the three gates regulate the flow of information into and out of the cell \ref{}.
+The cell remembers values over arbitrary time intervals, and the three gates regulate the flow of information into and out of the cell \cite{donahue_long-term_2016}.
 
 LSTM networks are well-suited to classifying, processing, and making predictions based on time series data since there can be lags of unknown duration between significant events in a time series.
-LSTMs are developed to deal with the vanishing gradient problems that can are encountered when training traditional RNNs \ref{}.
+LSTMs are developed to deal with the vanishing gradient problems that can are encountered when training traditional RNNs \cite{donahue_long-term_2016}.
 
@@
    \end{center}
-   \caption{Example of a LSTM unit.}
+   \caption{Example of a LSTM unit \cite{donahue_long-term_2016}.}
    \label{fig:reinforcementAgent}
@@
 Due to the superior performance and computationally tractable representation capability in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks \cite{kim_multimodal_2018}.
-Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated \ref{}.
+Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated \cite{kim_multimodal_2018}.
 Finally, the prediction is made on top of aggregated representation, usually with another neural network to capture the interactions between models and learn complex function mapping between input and output.
@@
    \begin{center}
-      \includegraphics[scale=0.6]{../media/multimodal_overview.png}
+      \includegraphics[scale=0.65]{../media/multimodal_overview.png}
    \end{center}
@@
    \end{center}
-   \caption{Example of QxQ quantile bin matrix used to calculate MTF.}
+   \caption{Example of QxQ quantile bin matrix used to calculate MTF \cite{wang_encoding_nodate}.}
    \label{fig:mtf_q}
@@
    \end{center}
-   \caption{CNN architecture for classifying images generated through MTF's.}
+   \caption{CNN architecture for classifying images generated through MTF.}
    \label{fig:mtf_cnn_model}
@@
 \section{Black-Box Model}
-Our black-box model is constructed of three different architectures: 1 LSTM and 2 CNN’s that are all concatenated at their respective dense layer.
+Our black-box model is constructed of three different architectures: 1 LSTM and 2 CNN’s that are all concatenated at their respective dense layer (Figure \ref{fig:multimodal_overview}).
 The three features we use are the amplitude/phase of each level of decomposition for HV.
 We chose three because we wanted to increase the dimensionality of our data in the event that one had more descriptive features than the others.
+
+Additionally, we tested our model using three levels of decomposition using the property VV as well as a combination of both VV and HV.
+We found no change in accuracy, therefore to limit the number of models we used, we only utilized the HV property.
 
@@
 
+In order to generate images using MTF, we have to select three properties, and we chose P\textsubscript{1} as the radial value of HV's amplitude value, P\textsubscript{2} as the radial value of VV's amplitude value, and P\textsubscript{3} as the absolute squared distance between every P\textsubscript{1} and P\textsubscript{2}.
+
+We then split those images using a randomized 70/30 split where 70\% of the data is our training set, and the other 30\% is our test set.
+We were able to achieve 60\% accuracy across all three of our datasets and show the features amongst the classes behave differently, as can be seen in Figure \ref{fig:whitebox_images}.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.5]{../media/whitebox_images.png}
+   \end{center}
+   \caption{Examples of images generated through MTF.}
+   \label{fig:whitebox_images}
+\end{figure}
+
 \chapter{Conclusion}

### 2020-03-17 paper/main.tex
added code and more citations

@@
 \usepackage{nameref}
+\usepackage{setspace}
+\usepackage{listings}
+
+\usepackage[toc,page]{appendix}
 
@@
 As time-series datasets become ever more popular, the need for deep learning models becomes that much greater.
-In our black-box model, we analyze deep learning models such as Convolutional Neural Networks, which have become increasingly popular and widely used in the last decade.
+In our black-box model, we analyze deep learning models such as Convolutional Neural Networks, which have become increasingly popular and widely used in the last decade  \cite{cui_multi-scale_2016}.
 However, their complex nature results in a lack of understanding of its inner workings \cite{wang_hybrid_2019}.
@@
 The method in which transmit antennae send and receive radio waves can be described through polarization.
-Two popular basis polarizations are horizontal linear, H, and vertical linear, V.
+Two popular basis polarizations are horizontal linear, H, and vertical linear, V \cite{oshea_convolutional_2016}.
 Our datasets contain data horizontally transmitted and vertically received (HV), as well as data vertically transmitted and vertically received (VV).
@@
 The Short-Time Fourier transform (STFT) is a Fourier related transform used to determine the sinusoidal frequency and phase content of local sections of a signal as it changes over time \cite{hill_uncertainty_nodate}.
-Computing STFT requires the signal to be divided into segments of equal length and then have the FT applied to that segment.
+Computing STFT requires the signal to be divided into segments of equal length and then have the FT applied to that segment \cite{allen_unified_1977}.
 This allows us to view the Fourier spectrum at a more granular level, which could potentially reveal different patterns amongst signals of different classes.
@@
 
-Wavelet transforms are similar to the FT in that they deconstruct a signal using representations of other signals.
+Wavelet transforms are similar to the FT in that they deconstruct a signal using representations of other signals \cite{misiti_wavelets_2013}.
 The key difference is that wavelet signals are finite in time and frequency as opposed to sine and cosine signals, which can carry on indefinitely \cite{strang_wavelets_1996}.
@@
    \end{center}
-   \caption{Example of different types of wavelets  \cite{strang_wavelets_1996}.}
+   \caption{Example of different types of wavelets \cite{lopez_development_2017}.}
    \label{fig:reinforcementAgent}
@@
 A signal is analyzed using different versions of dilated and translated basis functions called the mother wavelet  \cite{strang_wavelets_1996}.
-There are two types of wavelet transformations, discrete and continuous.
+There are two types of wavelet transformations, discrete and continuous \cite{imani_curve_2016}.
 In this thesis, we focus on discrete wavelet transformations, which use a discrete set of wavelet scales and translations to decompose the signal into a mutually orthogonal set of wavelets.
@@
 
-We use spline interpolation to resize each level of decomposition into a single array.
+We use spline interpolation to resize each level of decomposition into a single array because it allows to retain properties of our original data even in a higher dimension \cite{gregory_shape_1985}.
 We tested different sizes of interpolation from N to N/3, where N is the size of the largest discrete signal size, and found no reduction in performance.
@@
 
-Convolutional Neural Networks perform feature learning via non-linear transformations implemented as a series of layers.
+Convolutional Neural Networks perform feature learning via non-linear transformations implemented as a series of layers \cite{aqib_smarter_2019}.
 The input data is a multidimensional array, called a tensor.
@@
    \begin{center}
-      \includegraphics[scale=0.3]{../media/nn_highlight.jpeg}
+      \includegraphics[scale=0.5]{../media/cnn_example.png}
    \end{center}
-   \caption{Example of a Convolutional Neural Network \cite{yamashita_convolutional_2018}.}
+   \caption{Example of a Convolutional Neural Network \cite{alom_state---art_2019}.}
    \label{fig:cnn_example}
@@
 Due to the superior performance and computationally tractable representation capability in multiple domains such as visual, audio, and text, deep neural networks have gained tremendous popularity in multimodal learning tasks \cite{kim_multimodal_2018}.
-Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated \cite{kim_multimodal_2018}.
+Typically, domain-specific neural networks are used on different modalities to generate their representations, and the individual representations are merged or aggregated \cite{liu_learn_2018}.
 Finally, the prediction is made on top of aggregated representation, usually with another neural network to capture the interactions between models and learn complex function mapping between input and output.
@@
 
-To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set.
-We then performed three-fold cross-validation and took the average of our results as the outcome.
-
 \begin{figure}[h!]
@@
 
-\section{Dataset B}
+To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set.
+We then performed three-fold cross-validation and took the average of our results as the outcome.
+
+\section{Dataset C}
 
@@
 
+\begin{appendices}
+   \chapter{MTF Code}
+      \section{Main File}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible ]{../code/mtf/main.py}
+      \section{Calculating MTF}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/complex_signal.py}
+      \section{Helper functions}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/util.py}
+   \chapter{Hybrid Model Code}
+      \section{Main File}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/main.py}
+      \section{White-Box Model}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/white_box.py}
+      \section{Black-Box Model}
+         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/black_box.py}
+
+\end{appendices}
+
 \bibliography{../references/main}

### 2020-03-17 code/hybrid_model/black_box.py

@@ -1,0 +1,244 @@
+from keras.models import Sequential
+from keras.layers.normalization import BatchNormalization
+from keras.layers.convolutional import Conv1D, MaxPooling1D
+from keras.layers.core import Activation
+from keras.layers.core import Flatten
+from keras.layers.core import Dropout
+from keras.layers import Input, Dense, Concatenate
+from keras.layers import LSTM
+from keras.callbacks import CSVLogger
+from keras.layers import merge
+
+from keras.layers.core import Dense
+from keras import backend as K
+from keras.optimizers import Adam
+from keras.models import Model
+from sklearn.model_selection import StratifiedKFold
+from keras.utils import to_categorical
+import os
+from keras.callbacks import ModelCheckpoint
+
+import os, glob, shutil
+from scipy import signal
+from pywt import wavedec
+import pywt
+
+import scipy.ndimage
+
+def resize_(before, after_size=300):
+    '''
+    Function: resize_
+    Purpose: Resizes data using spline interpolation
+    '''
+    before_size = before.shape[0]
+    ratio = after_size / before_size
+    return scipy.ndimage.zoom(before, ratio)
+
+def processData(data1, data2, data3, train_split = .7, random=True):
+    '''
+        Function: processData
+        Purpose: Returns training and test data with the option of having them randomized
+    '''
+
+    #Generate array with numbers 0 - size of training data
+    #then shuffle the array
+    random_range = np.arange(data1.shape[0])
+    if random:
+        np.random.shuffle(random_range)
+
+    labels = np.asarray([0,1,2]* (data1.shape[0] // 3))
+    train_range = int(random_range.shape[0] * train_split)
+
+    x_train = data1[random_range[:train_range]]
+    x_train1 = data2[random_range[:train_range]]
+    x_train2 = data3[random_range[:train_range]]
+
+    y_train = labels[random_range[:train_range]]
+
+    x_test =  data1[random_range[train_range:]]
+    x_test1 =  data2[random_range[train_range:]]
+    x_test2 =  data3[random_range[train_range:]]
+
+    y_test = labels[random_range[train_range:]]
+
+    shuffleImageData()
+    return x_train, x_train1, x_train2, y_train, x_test, x_test1, x_test2, y_test
+
+def shuffleImageData(random_range, train_range):
+    '''
+    function: shuffleImageData
[…]

### 2020-03-17 code/hybrid_model/main.py

@@ -1,0 +1,48 @@
+import numpy as np
+import tensorflow as tf
+
+from black_box import blackBoxModel
+from white_box import whiteBoxModel
+
+from os import listdir
+from os.path import isfile, join
+
+def getListOfClassifiedImages(matched_predictions):
+    '''
+    Function: getListOfClassifiedImages
+    Purpose: Given indices of images, return a list of names corresponding to those indices
+    '''
+    test_data_path = './cnn_data/test'
+    image_files = [f for f in listdir(test_data_path) if isfile(join(test_data_path, f))]
+
+    selected_files = []
+    for i in range(len(image_files)):
+        if i in matched_predictions:
+            selected_files.append(image_files[i])
+
+    return selected_files
+def main():
+    '''
+    Function: main
+    Purpose: Call white-box & black-box model and get intersection of correctly classified signals
+    '''
+
+    signals_dataset_path = "../Channel_Impulse_Response"
+    image_dataset_path = "../data_whitebox"
+
+    test_results_black_box = blackBoxModel(signals_dataset_path)
+    test_results_white_box = whiteBoxModel(image_dataset_path)
+
+    # get index of predicted class
+    black_box_indexes = tf.argmax(test_results_black_box, axis=1)
+    black_box_indexes = black_box_indexes.eval()
+    white_box_indexes = tf.argmax(test_results_white_box, axis=1)
+    white_box_indexes = white_box_indexes.eval()
+
+    #getting the intersection of correctly classified items from black-box and white-box model
+    matched_predictions = np.where(white_box_indexes == black_box_indexes)
+
+    images = getListOfClassifiedImages(matched_predictions)
+
+if __name__== "__main__":
+  main()

### 2020-03-17 code/hybrid_model/white_box.py

@@ -1,0 +1,116 @@
+import keras
+from keras.models import Sequential
+from keras.layers import Dense, Activation, Dropout, Flatten, Conv2D, MaxPooling2D
+from keras.layers.normalization import BatchNormalization
+import numpy as np
+
+
+def whiteBoxModel(image_dataset_path):
+    '''
+    function: whiteBoxModel
+    Purpose: Main function that generates CNN model and evaluates on image dataset
+    '''
+
+    #define paramaters
+    training_data_size = 490
+    val_data_size = 210
+    batch_size = 32
+    epochs = 50
+    train_data_path = image_dataset_path + '/train'
+    val_data_path = image_dataset_path + '/val'
+    test_data_path = image_dataset_path + '/test'
+    img_size=224
+    
+    model = getCNN()
+    model.compile(loss=keras.losses.categorical_crossentropy, optimizer='adam', metrics=['accuracy'])
+
+    generator = ImageDataGenerator(
+        rotation_range=20,
+        width_shift_range=0.2,
+        height_shift_range=0.2,
+        horizontal_flip=True,
+        preprocessing_function=keras.applications.inception_v3.preprocess_input,
+        validation_split=0.3,
+    )
+
+    train_gen = generator.flow_from_directory(
+        train_data_path, 
+        batch_size=batch_size,    
+        subset='training', 
+        class_mode='categorical', 
+        target_size=(img_size, img_size)
+    )
+
+    val_gen = generator.flow_from_directory(
+        val_data_path,     
+        batch_size=batch_size,
+        subset='validation', 
+        class_mode='categorical', 
+        target_size=(img_size,img_size)
+    )
+
+    model.fit_generator(
+        generator=train_gen,
+        steps_per_epoch= training_data_size//batch_size ,
+        epochs= epochs,
+        validation_data=val_gen,
+        validation_steps= val_data_size // batch_size
+    )
+
+    datagen = ImageDataGenerator()
+
+    generator = datagen.flow_from_directory(
+                test_data_path,
+                target_size=(img_size, img_size),
+                batch_size=batch_size,
+                class_mode='categorical',
+                shuffle=True
+            ) 
+
+    return model.predict_generator(generator)
+
+def getCNN():
+    '''
+    function: getCNN
[…]

### 2020-03-17 code/main.py

@@ -1,212 +1,0 @@
-import numpy as np
-from scipy.io import loadmat
-import matplotlib.pyplot as plt
-import scipy.ndimage
-
-def resize_(before, after_size=5000):
-    before_size = before.shape[0]
-    ratio = after_size / before_size
-    return scipy.ndimage.zoom(before, ratio)
-
-
-import os, glob
-from scipy import signal
-from pywt import wavedec
-import pywt
-
-
-def generateData(path):
-    max_lev = 3
-
-    class_names = ['h_HV_SBR_1_discrete', 'h_HV_SBR_2_discrete', 'h_HV_SBR_3_discrete']
-    class_names_vv = ['h_VV_SBR_1_discrete', 'h_VV_SBR_2_discrete', 'h_VV_SBR_3_discrete']
-
-    data1=[]
-    data2=[]
-    data3 = []
-    num_classes = 3
-
-    c= 0
-    for _file in glob.glob(path):
-        filename = _file
-        file_contents = loadmat(filename)
-
-        c+=1
-        if c == 11:
-            break
-
-        for idx in range(num_classes):
-            fc = file_contents[class_names[idx]]
-            fc_vv = file_contents[class_names_vv[idx]]
-            
-            #normalize data
-            data = fc / np.linalg.norm(fc)
-            data_vv = fc_vv /np.linalg.norm(fc_vv)
-            
-            #Assign data to a variable so we can overwrite it after every level of decomposition
-            hv = data[:22]
-            vv = data_vv[:22]
-
-            for level in range(0, max_lev):
-
-                cA_hv, cD_hv = pywt.dwt(hv, 'db4', axis=0)
-                cA1_vv, cD_vv = pywt.dwt(vv, 'db4', axis=0)
-
-                hv = cD_hv
-                vv = cD_vv
-
-                if level == 0:
-                    data1.append(np.abs(cA_hv))
-
-                if level == 1:
-                    data2.append(np.abs(cA_hv))
-
-                if level == 2:
-                    data3.append(np.abs(cA_hv))  
-
-    data1 = np.asarray(data1)
-    data2 = np.asarray(data2)
-    data3 = np.asarray(data3)
-
-    train_test_split = .7
-
-    return processData(data1, data2, data3, train_test_split)
-
-
-'''
-    Function: processData
-    Purpose: Returns training and test data with the option of having them randomized
-'''
-def processData(data1, data2, data3, train_split = .7, random=True):
[…]

### 2020-03-17 code/mtf/complex_signal.py

@@ -1,0 +1,152 @@
+import numpy as np
+
+class ComplexSignal:
+  """
+  Class: ComplexSignal
+  Purpose: Process data with MTF
+  """
+  def __init__(self, time, theta, r, time_bins = 4, theta_bins = 4, r_bins = 4, image_size = 8, n_order = 4, _class=None):
+    n_samples, n_features = time.shape
+    self.image_size = image_size
+    self.window_size = n_features // image_size
+    self.time_series = [time, theta, r]
+    self.num_bins = [time_bins, theta_bins, r_bins]
+    self.sample = n_samples
+    self.features = n_features
+    self.mtf = None
+    self.remainder = n_features % image_size
+    self.n_order = n_order
+    self._class = _class
+
+  def _mtf(self, binned_ts, num_bins):
+    MTM = np.zeros((num_bins, num_bins))
+    
+    for i in range(self.n_order):
+      temp_MTM = np.zeros((num_bins, num_bins))
+
+      lagged_ts = np.vstack([binned_ts[:-1 - i], binned_ts[1 + i:]])
+
+      np.add.at(temp_MTM, tuple(map(tuple, lagged_ts)), 1)
+      temp_MTM *= (.4 ** (i+1))
+      MTM += temp_MTM
+
+    non_zero_rows = np.where(MTM.sum(axis=1) != 0)[0]
+    MTM = np.multiply(MTM[non_zero_rows][:, non_zero_rows].T,
+                      np.sum(MTM[non_zero_rows], axis=1)**(-1)).T
+
+    MTF = np.zeros((self.features, self.features))
+    list_values = [np.where(binned_ts == q) for q in non_zero_rows]
+
+    for i in range(non_zero_rows.size):
+        for j in range(non_zero_rows.size):
+            MTF[tuple(np.meshgrid(list_values[i], list_values[j]))] = MTM[i, j]
+    remainder = self.remainder
+    if remainder == 0:
+      return np.reshape(MTF,
+                        (self.image_size, self.window_size,
+                          self.image_size, self.window_size)
+                        ).mean(axis=(1, 3))
+    else:
+      self.window_size += 1
+      start, end, _ = segmentation(MTF.shape[0], self.window_size,
+                                    False, self.image_size)
+      AMTF = np.zeros((self.image_size, self.image_size))
+      for i in range(self.image_size):
+          for j in range(self.image_size):
[…]

### 2020-03-17 code/mtf/main.py

@@ -1,0 +1,98 @@
+import util
+import numpy as np
+from complex_signal import ComplexSignal
+import random
+from scipy.io import loadmat
+import os, glob
+from multiprocessing import Pool
+from keras.preprocessing.image import array_to_img
+import scipy.ndimage as sc
+from scipy.stats import entropy
+
+def resize_(before, after_size):
+    '''
+    Function: resize_
+    Purpose: Resizes data using spline interpolation
+    '''
+    before_size = before.shape[0]
+    ratio = after_size / before_size
+    return sc.zoom(before, ratio)
+
+def scalar_wrapper(arg):
+    '''
+    Function: scalar_wrapper
+    Purpose: Helper function to parallelize MTF
+    '''
+
+    arg.generateMTF()
+    return arg
+
+def main():
+
+    #define parameters for MTF
+    _IMAGE_SIZE = 300
+    N = 300
+    _CLASSES = 3
+    _TIME_BINS = 100
+    _THETA_BINS = 100
+    _R_BINS = 100
+    _ORDER = 1
+    max_level = 1
+    
+    data = ['h_HV_SBR_1_discrete', 'h_HV_SBR_2_discrete', 'h_HV_SBR_3_discrete']
+    data_vv = ['h_VV_SBR_1_discrete', 'h_VV_SBR_2_discrete', 'h_VV_SBR_3_discrete']
+    path_name = "./dataset"
+
+    for _file in glob.glob(path_name):
+
+      _data = []
+
+      filename = _file
+      file_contents = loadmat(filename)
+
+      data = np.asarray(data)
+      _CLASSES = data.shape[0]
+      i = 0
+
+      for _class in range(len(data)):
+          r, theta, time = [], [], []
+
+          #Read HV and VV property from data and normalize using L2 norm
+          _data_hv = file_contents[data[_class]][:,1]
+          _data_vv = file_contents[data_vv[_class]][:,1]
+          _data_hv /= np.linalg.norm(_data_hv)
+          _data_vv /= np.linalg.norm(_data_vv)
+
+          #Define features that will have MTF applied to them
+          r  = np.abs(_data_hv)
+          theta = np.abs(_data_vv)
+          time = ( np.abs(_data_hv) - np.abs(_data_vv) ) **2
+
+          #Resize data to allow for larger images to be created
+          r = resize_(np.abs(r), N)
+          theta = resize_(np.abs(theta), N)  
+          time  = resize_(np.abs(time), N)
+
[…]

### 2020-03-17 code/mtf/util.py

@@ -1,0 +1,56 @@
+import random
+import math
+import numpy as np
+
+def polar(z):
+    a= z.real
+    b= z.imag
+    r = math.hypot(a,b)
+    x = a / (math.sqrt(a**2 + b **2))
+    theta = math.acos(x)
+    return r,theta 
+
+def normalize(a):
+    return (a - np.min(a)) / (np.max(a) - np.min(a))
+
+def generateTimeSeries(N, i):
+    complex_numbers_time_series = []
+    time = 1
+    lower_bound = int(1 * i)
+    upper_bound = int(10* i)
+    for _ in range(N):
+        rand_complex_number = complex(random.randint(lower_bound, upper_bound), random.randint(lower_bound, upper_bound))
+        time = random.randint(int(time) + 1, int(time) + 10) + (10 - random.uniform(0., 10.))
+        time = time + (10 - random.uniform(0., 10.))        
+        
+        complex_numbers_time_series.append((time, rand_complex_number))
+
+    return complex_numbers_time_series
+
+
+def processRawMatrix(theta, time, r):
+    theta = np.asarray(normalize(theta)).reshape(1, -1)
+    r = np.asarray(normalize(r)).reshape(1, -1)
+    time = np.asarray(time).reshape(1, -1)
+
+    return theta, time, r
+
+def printHelper(theta, r, time, quantiles):
+    for i in range(len(theta)):
+        for j in range(len(theta[i])):
+            print("Data Point:", theta[i][j], r[i][j], time[i][j])
+            print("Quantile p,q,r:", quantile_bins[j])
+
+def printPolar(theta, r):
+    theta *= 2 * np.pi
+    ax = plt.subplot(projection='polar')
+    ax.set_rticks(r_bins.flatten())  # set radial ticks based on quantiles
+
+    ax.plot(theta, r, 'k.')
+
+    # ax.set_rticks([0.5, 1, 1.5, 2])  # less radial ticks
+    ax.set_rlabel_position(22.5)  # get radial labels away from plotted line
+    ax.grid(True)
+
+    # ax.set_title("Not scaled yet", va='bottom')
+    plt.show()

### 2020-03-17 paper/main.tex
typo

@@
 
-Given that this dataset is ten times larget than our other two datasets, we wanted to try different training and test splits, going as low as 8\% for training data and using the remaining 92\% as test data.
+Given that this dataset is ten times larger than our other two datasets, we wanted to try different training and test splits, going as low as 8\% for training data and using the remaining 92\% as test data.
 We ran the same number of cross-validations as our previous tests and found minimal difference in performance.

### 2020-03-20 paper/main.tex
added information about classes

@@
    \caption{Size of each dataset.}
-   \label{fig:stft_example}
+   \label{fig:dataset_info}
 \end{figure}
@@
 In this thesis, we focus on discrete data given that it is closer to datasets in real-world situations.
+
+\section{Dataset Classes}
+
+Dataset A and C were generated through similar methods and can be visually seen in Figure \ref{fig:dataset_ac}.
+There are three base stations, in which each base station corresponds to a class, sending information to a single receiver antenna.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.4]{../media/dataset_ac.png}
+   \end{center}
+   \caption{Class representation of how dataset A and C.}
+   \label{fig:dataset_ac}
+\end{figure}
+
+Dataset B differs from dataset A and C in how it transmits data.
+In \ref{fig:dataset_b}, we can see that there are three different antennas at the same station sending information to a single receiver station.
+The problem of classification becomes more challenging because they are very close together.
+
+\begin{figure}[h!]
+   \begin{center}
+      \includegraphics[scale=0.5]{../media/dataset_b.png}
+   \end{center}
+   \caption{Class representation of how dataset B.}
+   \label{fig:dataset_b}
+\end{figure}
 

### 2020-04-10 paper/main.tex
typo
new updates

@@
    By Ruben Rosales \\
+   B.S., 2016, California State University, Fullerton \\
    May 2020
@@
 Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification.
-These methods, however, fall short of allowing users to visualize patterns within their dataset.
+These methods, however, fall short of allowing users to visualize patterns, which can be a necessity for adoption in highly scrutinized industries.
 
@@
 \tableofcontents
-\listofillustrations
+\listoffigures
 
@@
 % This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
+\chapter{Related Work}
+\label{chap:relatedwork}
 
-\section{Contributions}
-We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model.
+The study of interpretable hybrid models has been worked on extensively in the past few years as deep learning has become increasingly popular.
+Scientists take different approaches as to what makes a hybrid model interpretable and apply it to different domains, such as computer vision or time series.
+For example, Wang \cite{wang_multilevel_2018}, focuses on creating a black-box model that utilizes wavelet decomposition for the classification of time-series that outperforms similar architectures.
+However, their white-box model focuses on analyzing the importance of their model's layers.
+This helps users identify the importance of wavelet decomposition but not necessarily understand their dataset.
+
+Additionally, other work we found tends to focus on non-time-series data or creating a general framework defining the trade-off between interpretability and black-box methods.
+For instance, \cite{wang_hybrid_2019} defines a set of rules to achieve a high level of classification with interpretability but does not give any competitive methods to achieve this.
+
+In order to fill the gap for a high performing interpretable model for time-series classification, we propose our hybrid model, which focuses on extensibility and robustness.
+We achieve this through a combination of deep learning models for classification as well as a unique white-box method to explain features within classes of a given dataset.
+
+% \section{Contributions}
+% We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model.
 Our final results show how robust our black-box model is with its competitive accuracy.
@@
 We analyzed our model on three radio signal datasets, which we will refer to as dataset A, B, and C throughout this thesis.
-These datasets were generated through Spatial Modulation which is a transmission technique that uses multiple antennae which maps a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}.
+These datasets were generated through Spatial Modulation, which is a transmission technique that uses multiple antennae to map a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}.
 The method in which transmit antennae send and receive radio waves can be described through polarization.
@@
    \end{center}
-   \caption{Class representation of how dataset A and C.}
+   \caption{Class representation of dataset A and C.}
    \label{fig:dataset_ac}
@@
    \end{center}
-   \caption{Class representation of how dataset B.}
+   \caption{Class representation of dataset B.}
    \label{fig:dataset_b}
@@
 % [explain hv, vv]
-
-\chapter{Related Work}
-\label{chap:relatedwork}
-
-The study of interpretable hybrid models has been worked on extensively in the past few years as deep learning has become increasingly popular.
-Scientists take different approaches as to what makes a hybrid model interpretable and apply it to different domains, such as computer vision or time series.
-For example, in \cite{wang_multilevel_2018}, they focus on creating a black-box model for the classification of time-series that outperforms similar architectures.
-However, their white-box model focuses on analyzing the importance of their model's layers.
-This helps users identify the importance of wavelet decomposition but not necessarily understand their dataset.
-
-Additionally, other work we found tends to focus on non-time-series data or creating a general framework defining the trade-off between interpretability and black-box methods.
-For instance, \cite{wang_hybrid_2019} defines a set of rules to achieve a high level of classification with interpretability but does not give any competitive methods to achieve this.
-
-In order to fill the gap for a high performing interpretable model for time-series classification, we propose our hybrid model, which focuses on extensibility and robustness.
-We achieve this through a combination of deep learning models for classification as well as a unique white-box method to explaining features within classes of a given dataset.
 
@@
 Each basis function is a complex value of a frequency, allowing us to view our data in the frequency domain as opposed to the amplitude domain.
+Our dataset are discrete in time domain so we focus on discrete fourier transforms.
+The discrete fourier transform of a signal x can be defined as
+
+\begin{equation} \label{eq1}
+   \begin{split}
+      X_{k} =  \sum_{n=0}^{N-1} x(n) e^{-\frac{i2\pi kn}{N}}
+   \end{split}
+\end{equation}
+
+where x(t\textsubscript{n}) is the input signal at time n, X(k) is the \textit{k}th spectral sample \cite{smith_mathematics_nodate}.
 
@@
 This allows us to view the Fourier spectrum at a more granular level, which could potentially reveal different patterns amongst signals of different classes.
+
+We can explain STFT through an example.
+Let x(n) be a signal, and let X\textsubscript{n}(e\textsuperscript{j$\omega\kappa$}) be the short-time Fourier transform of x(n) at time n and frequency $\omega$\textsubscript{$\kappa$} (Figure \ref{fig:stft_example}) \cite{allen_unified_1977}.
+
+\begin{figure}[h]
+   \begin{center}
+      \includegraphics[scale=0.6]{../media/stft_eq.png}
+   \end{center}
+   \caption{Example of short-time fourier transform formula \cite{allen_unified_1977}.}
+   \label{fig:stft_example}
+\end{figure}
 
@@
 
-\begin{figure}[h!]
-   \begin{center}
-      \includegraphics[scale=0.4]{../media/wavelet_families.png}
-   \end{center}
-   \caption{Example of different types of wavelets \cite{lopez_development_2017}.}
-   \label{fig:reinforcementAgent}
-\end{figure}
+% \begin{figure}[h!]
+%    \begin{center}
+%       \includegraphics[scale=0.4]{../media/wavelet_families.png}
+%    \end{center}
+%    \caption{Example of different types of wavelets \cite{lopez_development_2017}.}
+%    \label{fig:reinforcementAgent}
+% \end{figure}
 
@@
 \subsection{Each Level of Decomposition as a Dataset}
-Figure \ref{fig:multimodal_overview} depicts our model in which each level of decomposition is a unique dataset.
+Figure \ref{fig:wavelet_decomp} depicts our model in which each level of decomposition is a unique dataset.
 This method and the resampling method both achieved similar results, so we chose to move forward with this method.
@@
 Machine learning algorithms can be seen as learning a target function (f) that maps input data (X) to an output (Y).
+Then, f is used to make predictions on new input data.
 There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short-Term Memory networks (LSTM).
@@
 A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis.
-As seen in Figure X, it consists of an input and output layer as well as several hidden layers.
+As seen in Figure \ref{fig:cnn_example}, it consists of an input and output layer as well as several hidden layers.
 A convolution is an operation between a vector of weights w against an input x \cite{yamashita_convolutional_2018}.
@@
 
+To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set.
+We then performed three-fold cross-validation and took the average of our results as the outcome.
+As can be seen in Figure \ref{fig:final_results}, dataset A and C achieved similar accuracy, around 99\%, and dataset B averaged at around 91\% accuracy.
+
 \begin{figure}[h!]
@@
    \end{center}
-   \caption{Results}
+   \caption{Accuracy on test data for our black-box model.}
    \label{fig:final_results}
 \end{figure}
-
-To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set.
-We then performed three-fold cross-validation and took the average of our results as the outcome.
 
@@
 We then split those images using a randomized 70/30 split where 70\% of the data is our training set, and the other 30\% is our test set.
-We were able to achieve 60\% accuracy across all three of our datasets and show the features amongst the classes behave differently, as can be seen in Figure \ref{fig:whitebox_images}.
+Afterward, we trained our CNN (Figure \ref{fig:mtf_cnn_model}) on those images and were able to achieve 60\% accuracy across all three of our datasets and show the features amongst the classes behave differently, as can be seen in Figure \ref{fig:whitebox_images}.
 
@@
 
-\begin{appendices}
-   \chapter{MTF Code}
-      \section{Main File}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible ]{../code/mtf/main.py}
-      \section{Calculating MTF}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/complex_signal.py}
-      \section{Helper functions}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/util.py}
-   \chapter{Hybrid Model Code}
-      \section{Main File}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/main.py}
-      \section{White-Box Model}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/white_box.py}
-      \section{Black-Box Model}
-         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/black_box.py}
+% \begin{appendices}
+%    \chapter{MTF Code}
+%       \section{Main File}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible ]{../code/mtf/main.py}
+%       \section{Calculating MTF}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/complex_signal.py}
+%       \section{Helper functions}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/util.py}
+%    \chapter{Hybrid Model Code}
+%       \section{Main File}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/main.py}
+%       \section{White-Box Model}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/white_box.py}
+%       \section{Black-Box Model}
+%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/black_box.py}
 
-\end{appendices}
+% \end{appendices}
 

### 2020-04-13 paper/main.tex
new changes

@@
-\documentclass{turabian-thesis}
+\documentclass{turabian-thesis}[12pt]
 
@@
 \usepackage{listings}
+\usepackage{mathptmx}
+
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\l@section}
+  {\hfil}
+  {\leaders\hbox{\normalfont$\m@th\mkern \@dotsep mu\hbox{.}\mkern \@dotsep mu$}\hfill}
+  {}{}
+\makeatother
+\usepackage[english]{babel}
+\usepackage[font=bf, figurename=FIGURE, labelsep=period]{caption}
+
+\addto\captionsenglish{% Replace "english" with the language you use
+  \renewcommand{\contentsname}%
+    {TABLE OF CONTENTS}%
+}
+\usepackage{titlesec}
+
+% \usepackage{tocloft}
+% \renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
+
+\usepackage{tocbibind}
+\usepackage[titles]{tocloft}
+ \renewcommand \cftchapdotsep{1.5}
+
+ \renewcommand{\cfttoctitlefont}{\normalfont\Large}% Remove \bfseries from ToC title
+ \renewcommand{\cftsecfont}{}% Remove \bfseries from section titles in ToC
+ \renewcommand{\cftsecpagefont}{}% Remove \bfseries from section titles' page in ToC
+
+\let\chaptername\relax
+
+\renewcommand{\cftchapnumwidth}{6em}
+
+\addto{\captionsenglish}{\renewcommand{\bibname}{REFERENCES}}
+
+\addto\captionsenglish{\renewcommand{\chaptername}{CHAPTER}}
 
 \usepackage[toc,page]{appendix}
+
+% 1 inch margin
+\usepackage[margin=1in]{geometry}
 
@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{claim}{Claim}
-\newtheorem{definition}{Definition}
-\newtheorem{problem}{Problem}
-
-\newcounter{case}[section]
-\newenvironment{case}[1][]{\refstepcounter{case}\par\medskip
-   \textbf{Case~\thecase. #1} \rmfamily}{\medskip}
-
-\newcounter{subcase}[case]
-\newenvironment{subcase}[1][]{\refstepcounter{subcase}\par\medskip
-   \textbf{Case~\thecase.\thesubcase. #1} \rmfamily}{\medskip}
+\usepackage{fancyhdr}
+\fancypagestyle{headings}{%
+\cfoot{\thepage}
+}
+\pagestyle{headings}
+\fancyhead{}
+\renewcommand{\headrulewidth}{0pt}
 
@@
 
-   \textbf{Hybrid Model for Interpretable Time Series Analysis}
+   \textbf{HYBRID MODEL FOR INTERPRETABLE TIME SERIES ANALYSIS}
 
@@
 
-\chapter*{Abstract}
+\chapter*{ABSTRACT}
 
@@
 
-\addcontentsline{toc}{chapter}{Abstract}
-\chapter{Acknowledgments}
+\addcontentsline{toc}{chapter}{ABSTRACT}
+\chapter{ACKNOWLEDGMENTS}
 
@@
 
+\newpage
+% {\let\bfseries\mdseries% ...or \let\bfseries\relax
 \tableofcontents
+% }
+\newpage
+
+\renewcommand{\listfigurename}{LIST OF FIGURES}
 \listoffigures
@@
 \mainmatter
-\chapter{Introduction}
+\chapter{INTRODUCTION}
 \label{chap:introduction}
@@
 % This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
-\chapter{Related Work}
+\chapter{RELATED WORK}
 \label{chap:relatedwork}
@@
 
-\chapter{Data}
+\chapter{DATA}
 
@@
 
-\chapter{Preprocessing}
-\label{chap:preprocessing}
+\chapter{PREPROCESSING}
+\label{preprocessing}
 
@@
 
-\chapter{Models}
+\chapter{MODELS}
 \label{chap:models}
@@
 
-\chapter{Markov Transition Fields}
+\chapter{MARKOV TRANSITION FIELDS}
 \label{chap:mtf}
@@
 
-\chapter{Main Results}
+\chapter{MAIN RESULTS}
 \label{chap:results}
@@
 
-\chapter{Conclusion}
+\chapter{CONCLUSION}
 \label{chap:conclusion}
